### 2022-05-13 sample/tex/paper.tex
Add paper scaffold

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Fair Queuing Aware Congestion Control}

\author{\IEEEauthorblockN{Maximilian Bachl}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

% \begin{IEEEkeywords}
% component, formatting, style, styling, insert
% \end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}

\end{document}

### 2022-05-15 sample/tex/paper.tex
Draft of abstract

@@
 \documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
+% \IEEEoverridecommandlockouts
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
@@
 \usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{hyperref}
 \begin{document}
@@
 \author{\IEEEauthorblockN{Maximilian Bachl}
+\IEEEauthorblockA{{\scriptsize \url{https://github.com/muxamilian/fair-queuing-aware-congestion-control}}}
 % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+Fair queuing is becoming increasingly prevalent in the internet and has been shown to improve performance in many circumstances. 
+Performance could be improved even more if endpoints could detect the presence of fair queuing on a certain path and adjust their congestion control accordingly. 
+If fair queuing is detected, the congestion control would not have to take cross traffic into account, which allows for more flexibility. 
+In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path, while only adding small latency overhead and consuming no additional bandwidth. 
+When fair queuing is detected, a different congestion control is chosen, which results in reduced latency. 
+Unlike an algorithm proposed in a previous paper of us, the approach presented here does not only detect the presence of fair queuing once at flow startup but it does so continuously. 
 \end{abstract}

### 2022-05-15 picoquic/tonopah.c

@@ -49,7 +49,7 @@ typedef struct st_picoquic_tonopah_interval_info_t {
     struct st_picoquic_tonopah_interval_info_t* prev;
 } picoquic_tonopah_interval_info_t;
 
-#define INTERVALS_REQUIRED 8
+#define INTERVALS_REQUIRED 10
 
 picoquic_tonopah_interval_info_t* interval_list_first = NULL;
 picoquic_tonopah_interval_info_t* interval_list_last = NULL;
@@ -116,6 +116,7 @@ static void picoquic_tonopah_sim_enter_recovery(
     uint64_t current_time)
 {
     if (path1 != NULL && path2 != NULL) {
+        printf("Recovery: ");
         delete_info_list();
     }
     nr_state->ssthresh = nr_state->cwin / 2;
@@ -350,6 +351,7 @@ static void set_path(picoquic_cnx_t* cnx, picoquic_tonopah_sim_state_t* nr_state
                 // puts("Detected fq, lowering cw");
                 nr_state->ssthresh = (uint64_t) (((double) nr_state->cwin) * (7./8.));
                 nr_state->cwin = nr_state->ssthresh;
+                printf("FQ detected: ");
                 delete_info_list();
             }
             picoquic_tonopah_interval_info_t* new_interval = (picoquic_tonopah_interval_info_t*) malloc(sizeof(picoquic_tonopah_interval_info_t));

### 2022-05-15 sample/network_sim.py

@@ -1,3 +1,4 @@
+import glob
 import mininet
 import mininet.topo
 import mininet.net
@@ -54,7 +55,7 @@ class Opts:
     pass
 
 iperf = False
-max_time = 10
+max_time = 20
 congestion_control = "tonopah"
 
 def generate_tc_commands(if_name, with_delay=False):
@@ -66,17 +67,18 @@ def generate_tc_commands(if_name, with_delay=False):
         print("bdp", bdp)
     opt.buffer_size = None
     # opt.buffer_size = int(1. * math.ceil(bdp))
-    opt.buffer_size = 10
-    # opt.qdisc = 'fq'
-    opt.qdisc = 'pfifo'
+    # opt.buffer_size = 10
+    opt.qdisc = 'fq'
+    opt.qdisc = 'fq_codel'
+    # opt.qdisc = 'pfifo'
     opt.interface = if_name
 
     if with_delay:
-        if opt.qdisc != 'fq':
+        if opt.qdisc != 'pfifo':
             qdisc_string = f"{opt.qdisc}"
             if opt.buffer_size is not None: 
                  qdisc_string += f" limit {int(math.ceil(opt.buffer_size/2))}"
-        else:
+        elif opt.qdisc == 'fq':
             qdisc_string = f"{opt.qdisc} nopacing"
             if opt.buffer_size is not None: 
                  qdisc_string += f" flow_limit {int(math.ceil(opt.buffer_size/2))}"
@@ -176,6 +178,14 @@ if out:
     print("server out", out.decode("utf-8"))
 if err:
     print("server err", err.decode("utf-8"))
+
+files = list(filter(os.path.isfile, glob.glob("./*.client.qlog")))
+files.sort(key=lambda x: os.path.getmtime(x))
+most_recent_file = files[-1]
+
+if os.path.isfile("client.qlog"):
+    os.remove("client.qlog")
+os.symlink(most_recent_file, "client.qlog")
 
 mininet.cli.CLI(net)
 net.stop()

### 2022-05-15 sample/sample_server.c

@@ -431,9 +431,9 @@ int picoquic_sample_server(int server_port, const char* server_cert, const char*
             picoquic_set_default_congestion_algorithm(quic, picoquic_newreno_algorithm);
         }
 
-        picoquic_set_qlog(quic, qlog_dir);
-        picoquic_set_log_level(quic, 1);
-        picoquic_set_key_log_file_from_env(quic);
+        // picoquic_set_qlog(quic, qlog_dir);
+        // picoquic_set_log_level(quic, 1);
+        // picoquic_set_key_log_file_from_env(quic);
     }
 
     /* Wait for packets using the wait loop provided in the library.

### 2022-05-16 picoquic/frames.c
Sending the correct number of acks

@@ -2274,7 +2274,7 @@ void picoquic_compute_ack_gap_and_delay(picoquic_cnx_t* cnx, uint64_t rtt, uint6
 {
     if (cnx->congestion_alg->congestion_algorithm_number == PICOQUIC_CC_ALGO_NUMBER_TONOPAH && cnx->nb_paths == 2) {
         *ack_gap = 2;
-        *ack_delay_max = 5000ull;
+        *ack_delay_max = 10000ull;
     }
     return;
     
@@ -3596,6 +3596,9 @@ int picoquic_is_ack_needed_in_ctx(picoquic_cnx_t* cnx, picoquic_ack_context_t* a
 {
     int ret = 0;
 
+    // if (((struct sockaddr_in*) &(cnx->path[0]->local_addr))->sin_port != 4433 && !ack_ctx->act[is_opportunistic].ack_needed) {
+    //     puts("No ack needed");
+    // }
     if (ack_ctx->act[is_opportunistic].ack_needed) {
         if (pc != picoquic_packet_context_application || ack_ctx->act[is_opportunistic].ack_after_fin) {
             ret = 1;
@@ -3607,15 +3610,18 @@ int picoquic_is_ack_needed_in_ctx(picoquic_cnx_t* cnx, picoquic_ack_context_t* a
         else
         {
             uint64_t ack_gap = picoquic_ack_gap_override_if_needed(cnx, l_cid_sequence);
+            // printf("ack_gap: %lu, delay: %lu", ack_gap, cnx->ack_delay_remote);
 
             if (ack_ctx->act[is_opportunistic].highest_ack_sent + ack_gap <= picoquic_sack_list_last(&ack_ctx->sack_list) ||
                 ack_ctx->act[is_opportunistic].time_oldest_unack_packet_received + cnx->ack_delay_remote <= current_time) {
+                // puts("; doing something");
                 ret = 1;
             }
             else {
                 if (ack_ctx->act[is_opportunistic].time_oldest_unack_packet_received + cnx->ack_delay_remote < *next_wake_time) {
                     *next_wake_time = ack_ctx->act[is_opportunistic].time_oldest_unack_packet_received + cnx->ack_delay_remote;
                     SET_LAST_WAKE(cnx->quic, PICOQUIC_FRAME);
+                    // puts("; doing nothing");
                 }
             }
         }

### 2022-05-16 picoquic/sender.c

@@ -4477,9 +4477,16 @@ static int picoquic_select_next_path_mp(picoquic_cnx_t* cnx, uint64_t current_ti
     }
     // printf("i %d\n", i);
 
+    if (cnx->congestion_alg->congestion_algorithm_number == PICOQUIC_CC_ALGO_NUMBER_TONOPAH && cnx->nb_paths == 2) {
     /* Ensure that at most one path is marked as nominal ack path */
-    for (i += 1; i < cnx->nb_paths; i++) {
-        cnx->path[i]->is_nominal_ack_path = 0;
+        for (i += 1; i < cnx->nb_paths; i++) {
+            cnx->path[i]->is_nominal_ack_path = 1;
+        }
+    } else {
+        /* Ensure that at most one path is marked as nominal ack path */
+        for (i += 1; i < cnx->nb_paths; i++) {
+            cnx->path[i]->is_nominal_ack_path = 0;
+        }
     }
 
     if (i_min_rtt >= 0 || (cnx->congestion_alg->congestion_algorithm_number == PICOQUIC_CC_ALGO_NUMBER_TONOPAH && cnx->nb_paths == 2)) {

### 2022-05-16 picoquic/tonopah.c

@@ -166,6 +166,9 @@ void picoquic_tonopah_sim_notify(
 {
     uint64_t smoothed_rtt = path_x->smoothed_rtt;
     if (cnx->nb_paths > 1) {
+        if (cnx->nb_paths != 2) {
+            puts("Oh no, more paths than expected!");
+        }
         assert(cnx->nb_paths == 2);
         smoothed_rtt = (cnx->path[0]->smoothed_rtt + cnx->path[1]->smoothed_rtt)/2;
     }

### 2022-05-16 sample/network_sim.py

@@ -68,13 +68,14 @@ def generate_tc_commands(if_name, with_delay=False):
     opt.buffer_size = None
     # opt.buffer_size = int(1. * math.ceil(bdp))
     # opt.buffer_size = 10
-    opt.qdisc = 'fq'
-    opt.qdisc = 'fq_codel'
-    # opt.qdisc = 'pfifo'
+    # opt.qdisc = 'fq'
+    # opt.qdisc = 'fq_codel'
+    opt.qdisc = 'pfifo'
     opt.interface = if_name
 
+    qdisc_string = opt.qdisc
     if with_delay:
-        if opt.qdisc != 'pfifo':
+        if opt.qdisc == 'pfifo':
             qdisc_string = f"{opt.qdisc}"
             if opt.buffer_size is not None: 
                  qdisc_string += f" limit {int(math.ceil(opt.buffer_size/2))}"
@@ -82,6 +83,7 @@ def generate_tc_commands(if_name, with_delay=False):
             qdisc_string = f"{opt.qdisc} nopacing"
             if opt.buffer_size is not None: 
                  qdisc_string += f" flow_limit {int(math.ceil(opt.buffer_size/2))}"
+
     else:
         qdisc_string = opt.qdisc
 

### 2022-05-17 picoquic/tonopah.c
Improve algorithm in conjunction with fq_codel

@@ -49,7 +49,7 @@ typedef struct st_picoquic_tonopah_interval_info_t {
     struct st_picoquic_tonopah_interval_info_t* prev;
 } picoquic_tonopah_interval_info_t;
 
-#define INTERVALS_REQUIRED 10
+#define INTERVALS_REQUIRED 4
 
 picoquic_tonopah_interval_info_t* interval_list_first = NULL;
 picoquic_tonopah_interval_info_t* interval_list_last = NULL;
@@ -59,7 +59,8 @@ uint64_t updated_path1 = 0;
 
 uint64_t last_change = 0;
 
-double ratio = 9./16.;
+// double ratio = 9./16.;
+double ratio = 0.5625;
 
 picoquic_path_t* path1 = NULL;
 picoquic_path_t* path2 = NULL;
@@ -115,6 +116,10 @@ static void picoquic_tonopah_sim_enter_recovery(
     picoquic_congestion_notification_t notification,
     uint64_t current_time)
 {
+    if (nr_state->alg_state == picoquic_tonopah_alg_congestion_avoidance && get_interval_info_list_len(interval_list_first) == 0) {
+        puts("Tonopah: Packet lost but ignoring it");
+        return;
+    }
     if (path1 != NULL && path2 != NULL) {
         printf("Recovery: ");
         delete_info_list();
@@ -350,7 +355,7 @@ static void set_path(picoquic_cnx_t* cnx, picoquic_tonopah_sim_state_t* nr_state
             // memset(interval_list, 0, sizeof(interval_list));
             // interval_list->first_seq_num = picoquic_cc_get_sequence_number(cnx, actual_path);
             int detected_fq = aggregate_intervals(interval_list_last);
-            if (detected_fq) {
+            if (detected_fq && nr_state->alg_state == picoquic_tonopah_alg_congestion_avoidance) {
                 // puts("Detected fq, lowering cw");
                 nr_state->ssthresh = (uint64_t) (((double) nr_state->cwin) * (7./8.));
                 nr_state->cwin = nr_state->ssthresh;

### 2022-05-17 sample/network_sim.py

@@ -69,8 +69,8 @@ def generate_tc_commands(if_name, with_delay=False):
     # opt.buffer_size = int(1. * math.ceil(bdp))
     # opt.buffer_size = 10
     # opt.qdisc = 'fq'
-    # opt.qdisc = 'fq_codel'
-    opt.qdisc = 'pfifo'
+    opt.qdisc = 'fq_codel'
+    # opt.qdisc = 'pfifo'
     opt.interface = if_name
 
     qdisc_string = opt.qdisc

### 2022-05-27 picoquic/tonopah.c
Works across the whole range of network conditions

@@ -44,6 +44,7 @@ typedef struct st_picoquic_tonopah_interval_info_t {
     uint64_t dominant_path_id; // can be 1 or 2;
     uint8_t finished1;
     uint8_t finished2;
+    uint8_t dont_use;
     
     struct st_picoquic_tonopah_interval_info_t* next;
     struct st_picoquic_tonopah_interval_info_t* prev;
@@ -357,8 +358,8 @@ int aggregate_intervals(picoquic_tonopah_interval_info_t* list) {
             // } else {
             //     multiplier = 0.67;
             // }
-            // int sent_enough = bw_sent_ratio < 0.5 + 0.75 * (bw_sent_ratio-0.5);
-            int detected_fq = (observed_ratio < 0.5 + multiplier * (bw_sent_ratio-0.5));
+            int sent_enough = bw_sent_ratio > 0.5 + 0.75 * (ratio-0.5);
+            int detected_fq = (observed_ratio < 0.5 + multiplier * (bw_sent_ratio-0.5)) && sent_enough;
             // int detected_fq = (observed_ratio-0.5)/(bw_sent_ratio-0.5) < 0.5;
 
             // printf("detected_fq: %d, bw_dominant: %.2f, bw_submissive: %.2f, ratio: %.3f, ratio_sent: %.3f\n", detected_fq, bw_dominant, bw_submissive, observed_ratio, bw_sent_ratio);  
@@ -409,6 +410,7 @@ static void set_path(picoquic_cnx_t* cnx, picoquic_tonopah_sim_state_t* nr_state
             new_interval->prev = prev_element;
             if (interval_list_first == NULL) {
                 interval_list_first = new_interval;
+                new_interval->dont_use = 1;
             }
             interval_list_last = new_interval;
             if (prev_element != NULL) {
@@ -454,7 +456,9 @@ picoquic_tonopah_interval_info_t* find_right_interval(picoquic_cnx_t* cnx, picoq
                     if ((current_elem->prev->first_ack_time1 == 0)
                     || (current_elem->prev->last_ack_time1 == 0)
                     || (current_elem->prev->bytes_received1 == 0)
-                    || (current_elem->prev->first_seq_num1 == 0)) {
+                    || (current_elem->prev->first_seq_num1 == 0)
+                    // || (current_elem->prev->dont_use)
+                    ) {
                         current_elem->prev->finished1 = 0;
                     }
                     // assert(current_elem->prev->first_ack_time1 != 0);
[…]

### 2022-05-27 sample/network_sim.py

@@ -18,10 +18,7 @@ parser.add_argument('--qdisc', type=str, default='fq')
 
 args = parser.parse_args()
 
-
 mininet.clean.cleanup()
-subprocess.run('killall picoquic_sample'.split(' '))
-subprocess.run('killall iperf3'.split(' '))
 
 class MyTopo(mininet.topo.Topo):
     "Simple topology example."
@@ -62,201 +59,209 @@ class Opts:
     pass
 
 iperf = False
-max_time = 120
+max_time = 90
 congestion_control = "tonopah"
 
 results = []
-i = -1
 
 # # for delay in (100,):
 for delay in (10, 50, 100):
     results.append([])
     for rate in (10, 50, 100):
-# for delay, rate in ((10, 10),):
-    # for rate in (100,):
-        i += 1
-        # if i < 1: 
-        #     continue
-        print("delay", delay, "rate", rate)
+        results[-1].append([])
+        rep_counter = 0
+        while True:
+            subprocess.run('killall picoquic_sample'.split(' '))
+            subprocess.run('killall iperf3'.split(' '))
+            if rep_counter >= 5:
+                break
+            print("delay", delay, "rate", rate, "rep_counter", rep_counter)
 
-        opt = Opts()
-        # opt.qdisc = 'pfifo'
-        # opt.qdisc = 'fq_codel'
-        # opt.qdisc = 'fq'
-        opt.qdisc = args.qdisc
-        opt.delay = delay
-        opt.rate = rate
+            opt = Opts()
+            # opt.qdisc = 'pfifo'
+            # opt.qdisc = 'fq_codel'
+            # opt.qdisc = 'fq'
+            opt.qdisc = args.qdisc
+            opt.delay = delay
+            opt.rate = rate
 
-        def generate_tc_commands(if_name, with_delay=False):
-            global opt
-            bdp = (opt.delay/1000 * opt.rate*1000000)/(1500*8)
-            if with_delay:
-                print("bdp", bdp)
-            opt.buffer_size = None
-            # opt.buffer_size = int(1. * math.ceil(bdp))
-            # opt.buffer_size = 10
-            if opt.qdisc == 'pfifo' or opt.qdisc == 'fq':
-                # opt.buffer_size = bdp
-                opt.buffer_size = max(100, bdp)
-            opt.interface = if_name
+            def generate_tc_commands(if_name, with_delay=False):
+                global opt
[…]

### 2022-05-29 sample/tex/figures/fq_common_end.tex
Add paper introduction

%\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (1.75,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {interval};
%\draw[densely dotted] (1.75,0) -- (1.75,0.95);

\newcommand{\xaxislength}{1.8}

\draw[->, line width=0.05mm, -dot3-={0.5 cm}{1 rtt}, -dot3-={1.0 cm}{2 rtts}, -dot3-={1.5 cm}{3 rtts}] (0,0) -- ++(\xaxislength , 0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.65, 0.05) (time) {time};

\draw[->, line width=0.05mm] (0,0) -- ++(0,2.35);
\draw[draw=none, line width=0.05mm] (-0.05,-0.0) -- ++(0,1);
\draw[draw=none, line width=0.05mm] (0,-0.05) -- ++(0,1);

\node (red_label) at (0.1,1.95) [inner sep=0,fill=color1,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 2}] {};

\node (green_label) at (0.1,1.8) [inner sep=0,fill=color2,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 1}] {};

\end{tikzpicture}
\end{document}

### 2022-05-29 sample/tex/figures/fq_illustration_goodput_fq.tex
Add paper introduction

\input{fq_common_beginning.tex}

%\draw[draw,decorate,decoration={brace, mirror}] (2.5,0) -- ++(0.0,1) node [midway,right,align=left,font=\fontsize{3pt}{0}\selectfont] {buffer\\ size};
%
%\draw[dashed] (0,0) -- ++(2.5,0);
%\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal minimum\\ window (BDP)};
%\draw[->, line width=0.05mm] (optimal) -- ++(0,0.25);

\node at (\barwidth /2, \barheight /2) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=\barheight cm] {};
\node at (\barwidth /2, 2*\barheight) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=2*\barheight cm] {};

\node at (\barwidth /2 + \barwidth, \barheight) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=2*
\barheight cm] {};
\node at (\barwidth /2 + \barwidth, 4*\barheight) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=4*\barheight cm] {};

\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};

\draw[dashed] (0,1.2) -- ++(1.8,0);
\node[align=center,font=\fontsize{2.75pt}{0}\selectfont,inner sep=0mm] at (0.375, 1.35) (maximum) {link speed};
\draw[->, line width=0.05mm] (maximum) -- (0.375,1.2);

\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {receiving rate};

\input{fq_common_end.tex}

### 2022-05-29 sample/tex/figures/fq_illustration_goodput_no_fq.tex
Add paper introduction

--- 2022-05-29 sample/tex/figures/fq_illustration_goodput_fq.tex
@@
 
-\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
-\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 0.2) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=0.4 cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 0.8) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=0.8 cm] {};
 

### 2022-05-29 sample/tex/figures/fq_illustration_throughput.tex
Add paper introduction

--- 2022-05-29 sample/tex/figures/fq_illustration_goodput_fq.tex
@@
 
-\node at (\barwidth /2 + 2*\barwidth, 0.3) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
-\node at (\barwidth /2 + 2*\barwidth, 0.9) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=0.6cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 2*\barheight) [inner sep=0,fill=color2,rectangle,minimum width=\barwidth cm,minimum height=4*
+\barheight cm] {};
+\node at (\barwidth /2 + 2*\barwidth, 8*\barheight) [inner sep=0,fill=color1,rectangle,minimum width=\barwidth cm,minimum height=8*\barheight cm] {};
 
-\draw[dashed] (0,1.2) -- ++(1.8,0);
-\node[align=center,font=\fontsize{2.75pt}{0}\selectfont,inner sep=0mm] at (0.375, 1.35) (maximum) {link speed};
-\draw[->, line width=0.05mm] (maximum) -- (0.375,1.2);
+\node (green_label) at (0.1,1.8) [inner sep=0,fill=color2,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 1}] {};
 
-\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {receiving rate};
+\node[anchor=north west, align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.01, 2.3) (rate) {sending rate};
 

### 2022-05-29 sample/tex/paper.tex
Add paper introduction

@@
 \usepackage{hyperref}
+\usepackage{subfig}
 \begin{document}
@@
 If fair queuing is detected, the congestion control would not have to take cross traffic into account, which allows for more flexibility. 
-In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path, while only adding small latency overhead and consuming no additional bandwidth. 
+In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path. 
 When fair queuing is detected, a different congestion control is chosen, which results in reduced latency. 
@@
 \section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
 
-\section{Ease of Use}
+In our previous work we proposed a technique which determines the presence of fair queuing at flow startup \cite{bachl_detecting_2021}, which worked as follows: 
+If fair queuing is successfully detected at flow startup, a congestion control was used, which aimed to keep queuing delay low (delay-based congestion control). 
+While this delay-based congestion control achieved high throughput and low delay, it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,noauthor_cubic_nodate},
+similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}.
+This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus is only used if fair queuing is detected. 
+If it is detected that there is no fair queuing, our approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other network flows,
+but doesn't keep delay as low as our delay-based congestion control.
 
-\subsection{Maintaining the Integrity of the Specifications}
+While our previous approach had high detection accuracy (98\%), it also had some limitations:
+\begin{itemize}
+    \item It would \textbf{only} detect fair queuing at \textbf{flow startup}. 
+    But if the bottleneck link changes during a flow, it could be that the previous bottleneck had fair queuing while the new one doesn't. This wouldn't have been detected. 
+    \item It would detect fair queuing only after filling the queue at the bottleneck completely, \textbf{causing packet loss}. 
+\end{itemize}
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+We would thus like to have an approach which 
+\begin{itemize}
+    \item \textbf{continuously checks} for the presence or absence of fair queuing, not only at flow startup. 
+    \item \textbf{doesn't cause packet loss} while trying to determine if there is fair queuing or not. 
+    \item can be \textbf{transparently used} on top of any congestion control algorithm. 
+\end{itemize}
 
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+\section{Concept}
 
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
+% \begin{figure}
+% \centering
+% \subfloat[\footnotesize Sending rate]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
+% \label{fig:throughput}}
+% \subfloat[\footnotesize Receiving rate\\(no fair queuing)]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
+% \label{fig:goodput_no_fq}}
+% \subfloat[\footnotesize Receiving rate\\(fair queuing)]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
+% \label{fig:goodput_fq}}
+% \caption{An example illustrating our proposed flow startup mechanism.
+Figure \ref{fig:throughput} shows the sending rate, \ref{fig:goodput_no_fq} the receiving rate in case there's no fair queuing and \ref{fig:goodput_fq} the receiving rate if there is fair queuing.}
+% \label{fig:illustration}
+% \end{figure}
+
+\section{Evaluation}
+
+\section{Discussion}
+
+\bibliographystyle{ieeetr}
+\bibliography{fair-queuing-aware-congestion-control}
 

### 2022-05-29 picoquic/tonopah.c

@@ -52,7 +52,8 @@ typedef struct st_picoquic_tonopah_interval_info_t {
 
 #define INTERVALS_REQUIRED 4
 uint64_t minimum_interval = 50000;
-uint64_t maximum_interval = 100000;
+// uint64_t maximum_interval = 100000;
+uint64_t maximum_interval = 50000;
 
 picoquic_tonopah_interval_info_t* interval_list_first = NULL;
 picoquic_tonopah_interval_info_t* interval_list_last = NULL;
@@ -63,8 +64,8 @@ uint64_t updated_path1 = 0;
 uint64_t last_change = 0;
 
 // double ratio = 9./16.;
-// double ratio = 0.5625;
-double ratio = 0.6;
+double ratio = 0.5625;
+// double ratio = 0.6;
 
 picoquic_cnx_t * last_cnx = NULL;
 picoquic_path_t* path1 = NULL;

### 2022-05-29 sample/network_sim.py

@@ -59,15 +59,15 @@ class Opts:
     pass
 
 iperf = False
-max_time = 90
+max_time = 30
 congestion_control = "tonopah"
 
 results = []
 
-# # for delay in (100,):
-for delay in (10, 50, 100):
+# for delay in (10, 50, 100):
+for delay in (50, 50, 100):
     results.append([])
-    for rate in (10, 50, 100):
+    for rate in (50, 50, 100):
         results[-1].append([])
         rep_counter = 0
         while True:
@@ -152,20 +152,21 @@ for delay in (10, 50, 100):
             os.environ["MAX_TIME"] = str(max_time)
             os.environ["CONGESTION_CONTROL"] = congestion_control
 
-            # server_tcpdump_popen = h2.popen(f'tcpdump -s 100 -i h2-eth0 -w server.pcap (tcp || udp) and ip'.split(' '), **debug)
-            # client_tcpdump_popen = h1.popen(f'tcpdump -s 100 -i h1-eth0 -w client.pcap (tcp || udp) and ip'.split(' '), **debug)
-            server_tcpdump_popen = h2.popen(f'cat'.split(' '), **debug)
-            client_tcpdump_popen = h1.popen(f'cat'.split(' '), **debug)
+            server_tcpdump_popen = h2.popen(f'tcpdump -s 100 -i h2-eth0 -w server.pcap (tcp || udp) and ip'.split(' '), **debug)
+            client_tcpdump_popen = h1.popen(f'tcpdump -s 100 -i h1-eth0 -w client.pcap (tcp || udp) and ip'.split(' '), **debug)
+            # server_tcpdump_popen = h2.popen(f'cat'.split(' '), **debug)
+            # client_tcpdump_popen = h1.popen(f'cat'.split(' '), **debug)
 
             # server_popen = h2.popen('iperf3 -s -1'.split(' '), **debug)
             # time.sleep(1)
             # client_popen = h1.popen(f'iperf3 -c {h2.IP()}'.split(' '), **debug)
             server_popen = h2.popen(f'../picoquic_sample server 4433 ./ca-cert.pem ./server-key.pem ./server_files'.split(' '), **debug)
             if iperf:
-                iperf_server_popen = h1.popen(f'iperf3 -s'.split(' '), **debug)
+                iperf_server_popen = h1.popen(f'iperf3 -s'.split(' '), **{"stdout": None, "stderr": None})
             time.sleep(1)
             if iperf:
[…]

### 2022-05-29 sample/plot_bandwidth.py

@@ -13,7 +13,10 @@ sys.argv = sys.argv[:1]
 
 all_lines=[]
 for line in fileinput.input():
-	current_line = [float(item) for item in line.split(',')]
+	try:
+		current_line = [float(item) for item in line.split(',')]
+	except ValueError:
+		continue
 	if current_line[0] > args.to_time:
 		break
 	elif current_line[0] < args.from_time:
@@ -24,7 +27,14 @@ all_as_arrays = [np.array(item) for item in zip(*all_lines)]
 x = all_as_arrays[0]
 ys = all_as_arrays[1:]
 
-for y in ys:
+for i, y in enumerate(ys):
+	print("sum:", np.sum(y) / 1_000_000 * 8)
 	y = y / interval_len / 1_000_000 * 8
-	plt.plot(x, y)
+	plt.plot(x, y, alpha=0.67)
+
+plt.xlabel("Time (s)")
+plt.ylabel("Throughput (Mbit/s)")
+plt.ylim(bottom=0)
+plt.tight_layout()
+plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
 plt.show()

### 2022-05-29 sample/plot_rtt.py

@@ -2,22 +2,42 @@ import numpy as np
 import fileinput
 import matplotlib.pyplot as plt
 import json
-all_lines=[]
-for line in fileinput.input():
-    if "latest_rtt" not in line:
-        continue
-    line = line[:-2]
-    try:
-        parsed = json.loads(line)
-        t = float(parsed[0])
-        latest_rtt = float(parsed[4]['latest_rtt'])
-        all_lines.append((t, latest_rtt))
-    except Exception as e:
-        print("line", line)
-        raise e
+import sys
 
-x, y = [np.array(item) for item in zip(*all_lines)] 
-x = x / 1_000_000
-y = y / 1000.
-plt.plot(x, y)
+def add_plot(input_source):
+    all_lines=[]
+    for line in input_source:
+        if "latest_rtt" not in line:
+            continue
+        line = line[:-2]
+        try:
+            parsed = json.loads(line)
+            t = float(parsed[0])
+            latest_rtt = float(parsed[4]['latest_rtt'])
+            if latest_rtt > 0:
+                all_lines.append((t, latest_rtt))
+        except Exception as e:
+            print("line", line)
+            raise e
+
+    x, y = [np.array(item) for item in zip(*all_lines)] 
+    x = x / 1_000_000
+    y = y / 1000.
+    print("mean rtt:", np.mean(y))
+    plt.plot(x, y, alpha=0.67)
+
+if len(sys.argv) <= 1:
+    input_source = fileinput.input()
+    add_plot(input_source)
+else: 
+    with open(sys.argv[1]) as f:
+        input_source_1 = f.readlines()
+        add_plot(input_source_1)
+    with open(sys.argv[2]) as f:
+        input_source_2 = f.readlines()
+        add_plot(input_source_2)
+
+plt.xlabel("Time (s)")
+plt.ylabel("Delay (ms)")
+plt.ylim(bottom=0)
 plt.show()

### 2022-05-29 sample/sample_client.c

@@ -582,9 +582,9 @@ int picoquic_sample_client(char const * server_name, int server_port, char const
                 picoquic_set_default_congestion_algorithm(quic, picoquic_newreno_algorithm);
             }
 
-            // picoquic_set_key_log_file_from_env(quic);
-            // picoquic_set_qlog(quic, qlog_dir);
-            // picoquic_set_log_level(quic, 1);
+            picoquic_set_key_log_file_from_env(quic);
+            picoquic_set_qlog(quic, qlog_dir);
+            picoquic_set_log_level(quic, 1);
         }
     }
 

### 2022-06-01 sample/tex/paper.tex
Add to paper

@@
 \usepackage{subfig}
+\usepackage{tabularx}
 \begin{document}
@@
 % \end{IEEEkeywords}
+
+\section{Background}
+
+When different applications send packets on the internet, one application can send more than the other and thus take unfairly take a larger share of bandwidth. 
+This can result in unfairness and bad user experience.
+Several different approaches have been proposed to address this \cite{brown_future_2020,ware_beyond_2019} : One is to make sure every network flow is ``well-behaved'' (also known as TCP friendly)
+Another one is to enforce fairness at switches and routers, called ``fair queuing'' or ``flow queuing'' \cite{nagle_packet_1985}.
+
+While fair queuing was proposed decades ago, it only gained popularity in the last couple of years because of implementations in the Linux kernel \cite{dumazet_pkt_sched_2013,hoeiland-joergensen_flow_2018}.
+Applications can benefit from increasing deployment of fair queuing: It makes sure that not the most aggressive one wins. 
+It would be even better if applications could know if the connection they're sending on is managed by fair queuing.
+Then they could be sure that the can use a common congestion control mechanism, while not bothering or being bothered by other network flows.
+We proposed the first such approach in previous work \cite{bachl_detecting_2021} but our previous approach had some shortcomings upon which we improve in this paper.
+
+Our approach is a congestion control mechanism which also performs measurements to determine if there is fair queuing. 
+This approach of performing measurements in congestion control became popular in the last couple of years and was already followed by \cite{cardwell_bbr_2016,dong_pcc_2015,goyal_elasticity_2020,hayes_online_2020}.
 
@@
 
+\begin{figure}
+\centering
+\subfloat[\footnotesize Sending rate at the sender (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
+\label{fig:sender}}\\
+\subfloat[\footnotesize Receiving rate at the receiver (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
+\label{fig:receiver}}
+\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's fair queuing. 
+Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the flow that is sending more. 
+Thus, in the lower figure, the dominant flow and the non-dominant flow achieve approx.~the same receiving rate even though the dominant flow sends more (dominant flow).
+This is the effect of fair queuing.}
+\label{fig:illustration_fq}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfloat[\footnotesize Sending rate at the sender (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
+\label{fig:sender_pfifo}}\\
+\subfloat[\footnotesize Receiving rate at the receiver (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
+\label{fig:receiver_pfifo}}
+\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate in case there's no fair queuing. 
+Even though the sender sends too much and a queue builds up, still the flow that sends more data also has a higher receiving rate. 
+This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing.}
+\label{fig:illustration_no_fq}
+\end{figure}
+
 \section{Evaluation}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
+50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+\hline
+\end{tabularx}
+\caption{Detection accuracy in case there's \textbf{no fair queuing}. The average accuracy (true negative rate) is 100\%}
+\label{table:no_fq}
+\end{table}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\%\\ \hline
+50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+\hline
+\end{tabularx}
+\caption{Detection accuracy in case there is \textbf{fair queuing}. The average accuracy (true positive rate) is 96\%}
+\label{table:fq}
+\end{table}
 

### 2022-06-01 picoquic/bbr.c

@@ -754,6 +754,7 @@ void BBRExitStartupSeedBDP(picoquic_bbr_state_t* bbr_state, picoquic_path_t* pat
 
 void BBREnterProbeRTT(picoquic_bbr_state_t* bbr_state)
 {
+    // puts("ProbeRTT");
     bbr_state->state = picoquic_bbr_alg_probe_rtt;
     bbr_state->pacing_gain = 1.0;
     bbr_state->cwnd_gain = 1.0;
@@ -824,6 +825,7 @@ void BBRHandleProbeRTT(picoquic_bbr_state_t* bbr_state, picoquic_path_t * path_x
 
 void BBRCheckProbeRTT(picoquic_bbr_state_t* bbr_state, picoquic_path_t* path_x, uint64_t bytes_in_transit, uint64_t current_time)
 {
+    // printf("bbr_state->rt_prop_expired: %u\n", bbr_state->rt_prop_expired);
     if (bbr_state->state != picoquic_bbr_alg_probe_rtt &&
         bbr_state->rt_prop_expired &&
         !bbr_state->idle_restart) {

### 2022-06-01 sample/network_sim.py

@@ -60,7 +60,7 @@ class Opts:
 
 iperf = False
 max_time = 30
-congestion_control = "tonopah"
+congestion_control = "bbr"
 
 results = []
 
@@ -246,7 +246,7 @@ for delay in (50, 50, 100):
                     info.append((ts, value))
             print("info", info)
             correct_duration = 0.0
-            if len(info)-2 <= 0:
+            if len(info)-2 <= 0 and congestion_control == "tonopah":
                 print("Got too few results...")
                 continue
             for i in range(len(info)-2):

### 2022-06-01 sample/plot_bandwidth.py

@@ -27,14 +27,19 @@ all_as_arrays = [np.array(item) for item in zip(*all_lines)]
 x = all_as_arrays[0]
 ys = all_as_arrays[1:]
 
+print("default size:", plt.rcParamsDefault["figure.figsize"])
+plt.figure(figsize=(6.4, 4.8*0.67))
+
+labels = ['total', 'subflow 1', 'subflow 2']
 for i, y in enumerate(ys):
 	print("sum:", np.sum(y) / 1_000_000 * 8)
 	y = y / interval_len / 1_000_000 * 8
-	plt.plot(x, y, alpha=0.67)
+	plt.plot(x, y, alpha=0.67, label=labels[i])
 
 plt.xlabel("Time (s)")
 plt.ylabel("Throughput (Mbit/s)")
 plt.ylim(bottom=0)
+plt.legend(loc="lower right")
 plt.tight_layout()
 plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
 plt.show()

### 2022-06-01 sample/sample_client.c

@@ -578,6 +578,9 @@ int picoquic_sample_client(char const * server_name, int server_port, char const
             if (congestion_control != NULL && strcmp(congestion_control, "tonopah") == 0) {
                 picoquic_set_default_congestion_algorithm(quic, picoquic_tonopah_algorithm); 
             }
+            else if (congestion_control != NULL && strcmp(congestion_control, "bbr") == 0) {
+                picoquic_set_default_congestion_algorithm(quic, picoquic_bbr_algorithm); 
+            }
             else {
                 picoquic_set_default_congestion_algorithm(quic, picoquic_newreno_algorithm);
             }

### 2022-06-01 sample/sample_server.c

@@ -429,6 +429,9 @@ int picoquic_sample_server(int server_port, const char* server_cert, const char*
         if (congestion_control != NULL && strcmp(congestion_control, "tonopah") == 0) {
             picoquic_set_default_congestion_algorithm(quic, picoquic_tonopah_algorithm); 
         }
+        else if (congestion_control != NULL && strcmp(congestion_control, "bbr") == 0) {
+            picoquic_set_default_congestion_algorithm(quic, picoquic_bbr_algorithm); 
+        }
         else {
             picoquic_set_default_congestion_algorithm(quic, picoquic_newreno_algorithm);
         }

### 2022-06-10 picoquic/tonopah.c
Fix for fq_codel

@@ -52,8 +52,8 @@ typedef struct st_picoquic_tonopah_interval_info_t {
 
 #define INTERVALS_REQUIRED 4
 uint64_t minimum_interval = 50000;
-// uint64_t maximum_interval = 100000;
-uint64_t maximum_interval = 50000;
+uint64_t maximum_interval = 100000;
+// uint64_t maximum_interval = 50000;
 
 picoquic_tonopah_interval_info_t* interval_list_first = NULL;
 picoquic_tonopah_interval_info_t* interval_list_last = NULL;
@@ -64,8 +64,8 @@ uint64_t updated_path1 = 0;
 uint64_t last_change = 0;
 
 // double ratio = 9./16.;
-double ratio = 0.5625;
-// double ratio = 0.6;
+// double ratio = 0.5625;
+double ratio = 0.6;
 
 picoquic_cnx_t * last_cnx = NULL;
 picoquic_path_t* path1 = NULL;
@@ -198,23 +198,18 @@ void picoquic_tonopah_sim_notify(
         default: {
             uint64_t complete_delta = nb_bytes_acknowledged * path_x->send_mtu + nr_state->residual_ack;
             nr_state->residual_ack = complete_delta % nr_state->cwin;
-            // uint64_t current_smoothed_rtt = (cnx->path[0]->smoothed_rtt + cnx->path[1]->smoothed_rtt)/2;
-            // if (current_smoothed_rtt < minimum_interval || ((double) nr_state->cwin)/path_x->send_mtu <= 50) {
-            //     complete_delta *= 0.25*(((double)current_smoothed_rtt)/((double)minimum_interval));
-            // } 
-            // else 
-            // if (current_smoothed_rtt < minimum_interval || ((double) nr_state->cwin)/path_x->send_mtu <= 50) {
-            // if (((double) nr_state->cwin)/path_x->send_mtu <= 50) {
-            //     complete_delta *= 0.25*(((double)current_smoothed_rtt)/((double)minimum_interval));
-            // } 
-            nr_state->cwin += complete_delta / nr_state->cwin;
+            uint64_t current_smoothed_rtt = (cnx->path[0]->smoothed_rtt + cnx->path[1]->smoothed_rtt)/2;
+            double ratio = MIN((((double) current_smoothed_rtt) / ((double) minimum_interval)), 1.0);
+            // printf("ratio: %f\n", ratio);
+            nr_state->cwin += ratio * (((double) complete_delta) / ((double) nr_state->cwin));
[…]

### 2022-06-10 sample/network_sim.py

@@ -12,9 +12,12 @@ import os
 import subprocess
 import re
 import argparse
+import numpy as np
 
 parser = argparse.ArgumentParser()
 parser.add_argument('--qdisc', type=str, default='fq')
+parser.add_argument('--cc', type=str, default='tonopah')
+parser.add_argument('--iperf', action='store_true')
 
 args = parser.parse_args()
 
@@ -54,20 +57,24 @@ h2: mininet.node.Host = net.get('h2')
 s1: mininet.node.Host = net.get('s1')
 
 os.chdir(os.path.dirname(__file__))
+os.makedirs('logs', exist_ok=True)
 
 class Opts:
     pass
 
-iperf = False
-max_time = 30
-congestion_control = "bbr"
+max_time = 90
 
+bw_results = []
+delay_results = []
 results = []
 
-# for delay in (10, 50, 100):
-for delay in (50, 50, 100):
+for delay in (10, 50, 100):
+    bw_results.append([])
+    delay_results.append([])
     results.append([])
-    for rate in (50, 50, 100):
+    for rate in (10, 50, 100):
+        bw_results[-1].append([])
+        delay_results[-1].append([])
         results[-1].append([])
         rep_counter = 0
         while True:
@@ -103,18 +110,18 @@ for delay in (50, 50, 100):
                     if opt.qdisc == 'pfifo':
                         qdisc_string = f"{opt.qdisc}"
                         if opt.buffer_size is not None: 
-                            #  qdisc_string += f" limit {int(math.ceil(opt.buffer_size/2))}"
                             qdisc_string += f" limit {int(opt.buffer_size)}"
                     elif opt.qdisc == 'fq':
                         qdisc_string = f"{opt.qdisc} nopacing"
                         if opt.buffer_size is not None: 
-                            #  qdisc_string += f" flow_limit {int(math.ceil(opt.buffer_size/2))}"
                             qdisc_string += f" flow_limit {int(opt.buffer_size)}"
                     elif opt.qdisc == 'fq_codel':
-                        qdisc_string = f"{opt.qdisc} noecn"
-                        if opt.buffer_size is not None: 
-                            #  qdisc_string += f" flow_limit {int(math.ceil(opt.buffer_size/2))}"
-                            qdisc_string += f" flow_limit {int(opt.buffer_size)}"
+                        # min_delay = (1500 * 8) / (rate * 10**6) * 100
+                        # fq_codel_delay = int(max(min_delay*1000, 10))
[…]

### 2022-06-10 sample/picoquic_sample.h

@@ -40,8 +40,8 @@ extern "C" {
 
 #define PICOQUIC_SAMPLE_CLIENT_TICKET_STORE "sample_ticket_store.bin";
 #define PICOQUIC_SAMPLE_CLIENT_TOKEN_STORE "sample_token_store.bin";
-#define PICOQUIC_SAMPLE_CLIENT_QLOG_DIR ".";
-#define PICOQUIC_SAMPLE_SERVER_QLOG_DIR ".";
+#define PICOQUIC_SAMPLE_CLIENT_QLOG_DIR "./logs";
+#define PICOQUIC_SAMPLE_SERVER_QLOG_DIR "./logs";
 
 int picoquic_sample_client(char const* server_name, int server_port, char const* default_dir,
     int nb_files, char const** file_names);

### 2022-06-10 sample/plot_bandwidth.py

@@ -34,12 +34,12 @@ labels = ['total', 'subflow 1', 'subflow 2']
 for i, y in enumerate(ys):
 	print("sum:", np.sum(y) / 1_000_000 * 8)
 	y = y / interval_len / 1_000_000 * 8
-	plt.plot(x, y, alpha=0.67, label=labels[i])
+	# plt.plot(x, y, alpha=0.67, label=labels[i])
 
-plt.xlabel("Time (s)")
-plt.ylabel("Throughput (Mbit/s)")
-plt.ylim(bottom=0)
-plt.legend(loc="lower right")
-plt.tight_layout()
-plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
-plt.show()
+# plt.xlabel("Time (s)")
+# plt.ylabel("Throughput (Mbit/s)")
+# plt.ylim(bottom=0)
+# plt.legend(loc="lower right")
+# plt.tight_layout()
+# plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.show()

### 2022-06-10 sample/plot_rtt.py

@@ -24,7 +24,7 @@ def add_plot(input_source):
     x = x / 1_000_000
     y = y / 1000.
     print("mean rtt:", np.mean(y))
-    plt.plot(x, y, alpha=0.67)
+    # plt.plot(x, y, alpha=0.67)
 
 if len(sys.argv) <= 1:
     input_source = fileinput.input()
@@ -37,7 +37,7 @@ else:
         input_source_2 = f.readlines()
         add_plot(input_source_2)
 
-plt.xlabel("Time (s)")
-plt.ylabel("Delay (ms)")
-plt.ylim(bottom=0)
-plt.show()
+# plt.xlabel("Time (s)")
+# plt.ylabel("Delay (ms)")
+# plt.ylim(bottom=0)
+# plt.show()

### 2022-06-11 sample/tex/paper.tex
Add to paper

@@
 \usepackage{tabularx}
+\renewcommand{\thefootnote}{\arabic{footnote}}
+\renewcommand{\thempfootnote}{\arabic{mpfootnote}}
 \begin{document}
@@
 If fair queuing is successfully detected at flow startup, a congestion control was used, which aimed to keep queuing delay low (delay-based congestion control). 
-While this delay-based congestion control achieved high throughput and low delay, it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,noauthor_cubic_nodate},
+While this delay-based congestion control achieved high throughput and low delay, it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,ha_cubic_2008},
 similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}.
@@
 100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-\hline
 \end{tabularx}
-\caption{Detection accuracy in case there's \textbf{no fair queuing}. The average accuracy (true negative rate) is 100\%}
+\caption{Detection accuracy in case there's \textbf{no fair queuing}.  The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
 \label{table:no_fq}
@@
 100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+\end{tabularx}
+\caption{Detection accuracy in case there is \textbf{fair queuing}.  The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+\label{table:fq}
+\end{table}
+
+\subsection{Throughput and delay of Tonopah}
+
+\begin{table}
+\begin{minipage}{\columnwidth}
+\centering
+\begin{tabular}{| r | r | r |}
 \hline
+& Link utilization & Queuing delay \\ \hline
+Newreno & 95.6\% & 10.5 ms \\ \hline
+Tonopah & 94.3\% & 33.6 ms \\ \hline
+% p-value\protect\footnote{Using Welch's t-test} & 0.86 & $1.9 \times 10^{-6}$ \\ \hline
+% $\frac{\text{Newreno}}{\text{Tonopah}}$ & 103\% & 321\% \\ \hline
+\end{tabular}
+\caption{Newreno's link utilization is 103\% Tonopah's.
+But the \textbf{queuing delay} of Newreno is \textbf{321\%} Tonopah's.
+This means that Tonopah can deliver the same throughput while cause a lot less queuing delay.
+The differences in the throughput of Newreno and Tonopah are not significant (p-value\protect\footnote{Using Welch's t-test} $> 0.05$) while the differences in \textbf{queuing delay} are \textbf{highly significant} (p-value $< 10^{-5}$).}
+\label{table:throughput_and_delay}
+\end{minipage}
+\end{table}
+
+\subsection{Cross-traffic}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\%\\ \hline
+50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+100ms & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
 \end{tabularx}
-\caption{Detection accuracy in case there is \textbf{fair queuing}. The average accuracy (true positive rate) is 96\%}
-\label{table:fq}
+\caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}.
+The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+\label{table:no_fq_crosstraffic}
+\end{table}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
+50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+\end{tabularx}
+\caption{Detection accuracy in case there is \textbf{fair queuing} under the presence of \textbf{cross-traffic}.
+The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+\label{table:fq_crosstraffic}
+\end{table}
+
+\subsection{Other variants of fair queuing}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 85\% \newline 1st quart.: 84\% \newline 3rd quart.: 85\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
+50ms & Median: 91\% \newline 1st quart.: 88\% \newline 3rd quart.: 92\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 98\% \newline 1st quart.: 97\% \newline 3rd quart.: 99\% \\ \hline
+100ms & Median: 9\% \newline 1st quart.: 7\% \newline 3rd quart.: 13\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 64\% \newline 1st quart.: 0\% \newline 3rd quart.: 75\% \\ \hline
+\end{tabularx}
+\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing}.
+The overall median accuracy is 95\%, the first quartile is 83\% and the third quartile is 100\%.}
+\label{table:fq_codel}
 \end{table}

### 2022-06-11 sample/tex/paper.tex
Paper fix

@@
 This means that Tonopah can deliver the same throughput while cause a lot less queuing delay.
-The differences in the throughput of Newreno and Tonopah are not significant (p-value\protect\footnote{Using Welch's t-test} $> 0.05$) while the differences in \textbf{queuing delay} are \textbf{highly significant} (p-value $< 10^{-5}$).}
+The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value\protect\footnote{Using Welch's t-test} $< 10^{-5}$).}
 \label{table:throughput_and_delay}

### 2022-06-12 sample/tex/paper.tex
Add concept section to the paper

@@
 \usepackage{tabularx}
+\usepackage{dblfloatfix}
 \renewcommand{\thefootnote}{\arabic{footnote}}
@@
 \maketitle
+\thispagestyle{plain}
+\pagestyle{plain}
 
@@
 
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize \textbf{Sending} rate (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
+\label{fig:sender}}\\
+\subfloat[\footnotesize \textbf{Receiving} rate (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
+\label{fig:receiver}}
+\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's \textbf{fair queuing}. 
+Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the flow that is sending more. 
+Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more (dominant subflow).
+This is the effect of fair queuing.}
+\label{fig:illustration_fq}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize \textbf{Sending} rate (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
+\label{fig:sender_pfifo}}\\
+\subfloat[\footnotesize \textbf{Receiving} rate (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
+\label{fig:receiver_pfifo}}
+\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate in case there's \textbf{no fair queuing}. 
+Even though the sender sends too much and a queue builds up, still the flow that sends more data also has a higher receiving rate. 
+This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing.}
+\label{fig:illustration_no_fq}
+\end{figure}
+
 \section{Background}
 
-When different applications send packets on the internet, one application can send more than the other and thus take unfairly take a larger share of bandwidth. 
+When different applications send packets on the internet, one application can send more than the other and thus unfairly take a larger share of bandwidth. 
 This can result in unfairness and bad user experience.
@@
 
-% \begin{figure}
-% \centering
-% \subfloat[\footnotesize Sending rate]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_throughput"}.pdf}
-% \label{fig:throughput}}
-% \subfloat[\footnotesize Receiving rate\\(no fair queuing)]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_no_fq"}.pdf}
-% \label{fig:goodput_no_fq}}
-% \subfloat[\footnotesize Receiving rate\\(fair queuing)]{\includegraphics[width=0.33\columnwidth]{{"figures/fq_illustration_goodput_fq"}.pdf}
-% \label{fig:goodput_fq}}
-% \caption{An example illustrating our proposed flow startup mechanism.
-Figure \ref{fig:throughput} shows the sending rate, \ref{fig:goodput_no_fq} the receiving rate in case there's no fair queuing and \ref{fig:goodput_fq} the receiving rate if there is fair queuing.}
-% \label{fig:illustration}
-% \end{figure}
+The core concept of our approach, which we call \textbf{Tonopah}, is that a network flow is separated in \textbf{two subflows}. 
+\textbf{Alternatingly}, one flow sends more (\textbf{dominant subflow}), while the other one sends less (\textbf{non-dominant subflow}).
+After a certain time interval the dominant subflow becomes the non-dominant one and vice versa.
+When the total sending rate reaches the link capacity and fair queuing is present, \textbf{fair queuing} is going to \textbf{limit} the throughput of the \textbf{dominant subflow} (\autoref{fig:illustration_fq}).
+This can be measured by observing the \textit{delivery rate}
+\footnote{The delivery rate (a concept popularized by BBR \cite{cardwell_bbr_2016}) can be measured by observing the rate at which acknowledgement packets are sent back to the sender by the receiver.
+If the sender sends more than the link capacity, the delivery rate is never going to exceed the link capacity since the receiver cannot receive packets faster than the link capacity allows. 
+}
+of both subflows.
+If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}), which can be considered unfair.
+By alternating which subflow is dominant and which one is non-dominant, we prevent the build-up of a standing queue at the bottleneck:
+When the dominant subflow builds up a queue at the bottleneck link, it is going to be the non-dominant subflow in the next time interval and then it is going to send less and the queue is going to be reduced.
 
-\begin{figure}
-\centering
-\subfloat[\footnotesize Sending rate at the sender (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
-\label{fig:sender}}\\
-\subfloat[\footnotesize Receiving rate at the receiver (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
-\label{fig:receiver}}
-\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's fair queuing. 
-Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the flow that is sending more. 
-Thus, in the lower figure, the dominant flow and the non-dominant flow achieve approx.~the same receiving rate even though the dominant flow sends more (dominant flow).
-This is the effect of fair queuing.}
-\label{fig:illustration_fq}
-\end{figure}
+The congestion window of the flow applies to both subflows together.
+This means that while always the dominant subflow sends more than the non-dominant one, in sum they always send the same amount of data, as allowed by the congestion window.
 
-\begin{figure}
-\centering
-\subfloat[\footnotesize Sending rate at the sender (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
-\label{fig:sender_pfifo}}\\
-\subfloat[\footnotesize Receiving rate at the receiver (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
-\label{fig:receiver_pfifo}}
-\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate in case there's no fair queuing. 
-Even though the sender sends too much and a queue builds up, still the flow that sends more data also has a higher receiving rate. 
-This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing.}
-\label{fig:illustration_no_fq}
-\end{figure}
+Thanks to the use of Multipath QUIC \cite{liu_multipath_2022} the application does not see any difference because all data are recombined to one flow by Multipath QUIC.
+
+\subsection{Details}
+
+The dominant subflow receives $\frac{6}{10}$ of the total allowed sending rate, while the non-dominant one receives $\frac{4}{10}$. 
+The dominant subflow and the non-dominant subflow alternate every $\max{50, \min{100, \textit{rtt}}}$ milliseconds. 
+This time period is also called an \textbf{interval}. To determine the presence of fair queuing, the data of the last four intervals are evaluated. 
+We look at more than one interval because this reduces the effect of measurement noise. We found four intervals to be a suitable number.
+
+We determine that fair queuing is present if the receiving rate was more fair than not fair. 
+It would be completely fair if both flows get $\frac{1}{2}$ of the total throughput. It would be completely unfair if the dominant subflow got $\frac{6}{10}$ and the non-dominant one $\frac{4}{10}$. 
+If the dominant one got a fraction of 0.549 of the total throughput, it would be just a little bit more fair than unfair. In this case we would say that there is fair queuing. 
+If, on the other side, the dominant subflow got a fraction of 0.551 of the total throughput, we would say that there is no fair queuing.
+
+Another safeguard against measurement noise is that we only take a decision regarding fair queuing if the dominant flow actually really sent more data than the non-dominant one. 
+Specifically, if the dominant flow sends less than 57.5\%, we don't take a decision because the pacing apparently wasn't accurate enough. 
+This case only occurs when the sending rate is very low ($<$ 10\,Mbit/s) and thus pacing in inaccurate.
+
+Both subflows are controlled by the same congestion window. The differences in sending rates are implemented by using pacing. 
+The basis of the implementation of Tonopah is \textit{Picoquic}\footnote{\url{https://github.com/private-octopus/picoquic/tree/master/picoquic}}. 
+Picoquic includes support for Multipath QUIC and pacing, which we need for the implementation of Tonopah. Also, it implements congestion control algorithms. 
+We base Tonopah on the Newreno implementation of Picquic.
 

### 2022-06-12 sample/plot_bandwidth.py

@@ -34,12 +34,21 @@ labels = ['total', 'subflow 1', 'subflow 2']
 for i, y in enumerate(ys):
 	print("sum:", np.sum(y) / 1_000_000 * 8)
 	y = y / interval_len / 1_000_000 * 8
-	# plt.plot(x, y, alpha=0.67, label=labels[i])
+	plt.plot(x, y, alpha=0.67, label=labels[i])
 
-# plt.xlabel("Time (s)")
-# plt.ylabel("Throughput (Mbit/s)")
-# plt.ylim(bottom=0)
-# plt.legend(loc="lower right")
-# plt.tight_layout()
-# plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
-# plt.show()
+
+# plt.arrow(12.05, 10, 0, 12, color="black", edgecolor="black")
+plt.annotate("fair queuing", xy=(12.05, 23), xytext=(12.05, 10),
+            arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
+
+# plt.arrow(12.05, 10, 0, 12, color="black", edgecolor="black")
+plt.annotate("fair queuing", xy=(13.4, 23), xytext=(13.4, 10),
+            arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
+
+plt.xlabel("Time (s)")
+plt.ylabel("Throughput (Mbit/s)")
+plt.ylim(bottom=0)
+plt.legend(loc="lower left")
+plt.tight_layout()
+plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
+plt.show()

### 2022-06-12 sample/plot_rtt.py

@@ -24,7 +24,7 @@ def add_plot(input_source):
     x = x / 1_000_000
     y = y / 1000.
     print("mean rtt:", np.mean(y))
-    # plt.plot(x, y, alpha=0.67)
+    plt.plot(x, y, alpha=0.67)
 
 if len(sys.argv) <= 1:
     input_source = fileinput.input()
@@ -37,7 +37,7 @@ else:
         input_source_2 = f.readlines()
         add_plot(input_source_2)
 
-# plt.xlabel("Time (s)")
-# plt.ylabel("Delay (ms)")
-# plt.ylim(bottom=0)
-# plt.show()
+plt.xlabel("Time (s)")
+plt.ylabel("Delay (ms)")
+plt.ylim(bottom=0)
+plt.show()

### 2022-06-13 sample/tex/paper.tex
Add to results section

@@
 Unlike an algorithm proposed in a previous paper of us, the approach presented here does not only detect the presence of fair queuing once at flow startup but it does so continuously. 
+We show that by adding fair queuing detection to an existing congestion control algorithm, queuing delay can be reduced by more than two thirds. 
 \end{abstract}
@@
 
-The core concept of our approach, which we call \textbf{Tonopah}, is that a network flow is separated in \textbf{two subflows}. 
+The core concept of our approach, which we call \textbf{\textit{Tonopah}}, is that a network flow is separated in \textbf{two subflows}. 
 \textbf{Alternatingly}, one flow sends more (\textbf{dominant subflow}), while the other one sends less (\textbf{non-dominant subflow}).
@@
 
+An interesting aspect of fair queuing detection is that the presence of fair queuing can be interpreted as a \textit{congestion signal}: 
+Only if there is congestion, fair queuing can be detected. If the link is underutilized, flows don't ``fight'' for bandwidh and thus fairness doesn't need to be enforced at the bottleneck link. 
+Thus, in this paper, we define a new congestion signal: The presence of fair queuing. Other already known congestion signals are, for example: 
+packet loss; queue length, change of queue length, Explicit Congestion Notification \cite{mathis_relentless_2009,hutchison_revisiting_2011}.
+
 \subsection{Details}
@@
 The dominant subflow receives $\frac{6}{10}$ of the total allowed sending rate, while the non-dominant one receives $\frac{4}{10}$. 
-The dominant subflow and the non-dominant subflow alternate every $\max{50, \min{100, \textit{rtt}}}$ milliseconds. 
+The dominant subflow and the non-dominant subflow alternate every $\max{(50, \min{(100, \textit{rtt})})}$ milliseconds. 
 This time period is also called an \textbf{interval}. To determine the presence of fair queuing, the data of the last four intervals are evaluated. 
@@
 Picoquic includes support for Multipath QUIC and pacing, which we need for the implementation of Tonopah. Also, it implements congestion control algorithms. 
-We base Tonopah on the Newreno implementation of Picquic.
+We base Tonopah on the NewReno implementation of Picquic. When fair queuing is detected, the congestion window is reduced by $\frac{1}{8}$. 
+This worked well for our proof-of-concept implementation but one could also specify any other behavior in case fair queuing is detected or even do nothing at all.
 
 \section{Evaluation}
+
+Tonopah was implemented on top of Picquic in C. It was added on top of the NewReno congestion control algorithm but it could easily be added on top of other 
+congestion control algorithms as well. By basing our implementation on Picoquic, Tonopah also supports Explicit Congestion Notification.
+
+We evaluated Tonopah on an Apple MacBook Air from 2011 with an Intel Core i5-2557M CPU at 1.70GHz. 
+By choosing old hardware for experimentation, we can show that our algorithm doesn't use excessive compute and can also run well on weak hardware.
 
@@
 
-\subsection{Throughput and delay of Tonopah}
+\subsection{Throughput and Queuing Delay of Tonopah}
 
@@
 & Link utilization & Queuing delay \\ \hline
-Newreno & 95.6\% & 10.5 ms \\ \hline
+NewReno & 95.6\% & 10.5 ms \\ \hline
 Tonopah & 94.3\% & 33.6 ms \\ \hline
 % p-value\protect\footnote{Using Welch's t-test} & 0.86 & $1.9 \times 10^{-6}$ \\ \hline
-% $\frac{\text{Newreno}}{\text{Tonopah}}$ & 103\% & 321\% \\ \hline
+% $\frac{\text{NewReno}}{\text{Tonopah}}$ & 103\% & 321\% \\ \hline
 \end{tabular}
-\caption{Newreno's link utilization is 103\% Tonopah's.
-But the \textbf{queuing delay} of Newreno is \textbf{321\%} Tonopah's.
+\caption{NewReno's link utilization is 103\% Tonopah's.
+But the \textbf{queuing delay} of NewReno is \textbf{321\%} Tonopah's.
 This means that Tonopah can deliver the same throughput while cause a lot less queuing delay.
@@
 \end{table}
+
+While the goal of this paper is to implement a mechanism to detect fair queuing in an online fashion, 
+we also want to show that fair queuing detection can be used to lower the queuing delay drastically while not impairing throughput. 
+\autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than two thirds while keeping throughput unchanged, when it is implemented on top on NewReno
 
@@
 
+We also evaluated the performance of our fair queuing detection on a bottleneck managed by fq\_codel \cite{hoeiland-joergensen_flow_2018}. 
+We chose a default target queuing delay of 10\,ms following Apple's implementation\footnote{
+    \url{https://github.com/apple/darwin-xnu/blob/main/bsd/net/if.h\#L262}
+} because we argue that Apple probably spent a considerable amount of time fine-tuning their implementation and 
+came to the conclusion that 10\,ms work best as the default target delay.
+
 \begin{table}

### 2022-06-15 sample/tex/paper.tex
Add almost all results

@@
 Unlike an algorithm proposed in a previous paper of us, the approach presented here does not only detect the presence of fair queuing once at flow startup but it does so continuously. 
-We show that by adding fair queuing detection to an existing congestion control algorithm, queuing delay can be reduced by more than two thirds. 
+% We show that by adding fair queuing detection to an existing congestion control algorithm, queuing delay can be reduced by more than two thirds. 
 \end{abstract}
@@
 }
-of both subflows.
-If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}), which can be considered unfair.
+of both subflows. If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}), which can be considered unfair. 
+In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection.
+
 By alternating which subflow is dominant and which one is non-dominant, we prevent the build-up of a standing queue at the bottleneck: 
@@
 
+We emulate the network using mininet and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing in \textit{fq} for fair queuing.
+
+In the following evaluation we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
+If in the evaluation scenario there is fair queuing, and Tonopah detects fair queuing for 80\,s and fails to detect it for 10\,s we 
+state the accuracy as $\frac{80}{90} = 89\%$. 
 \begin{table}
@@
 
+\autoref{table:no_fq} shows that if there is no fair queuing, Tonopah detects this correctly virtually always. 
+If there is fair queuing (\autoref{table:fq}), Tonopah detects this correctly very often but not in every case.
+
 \subsection{Throughput and Queuing Delay of Tonopah}
@@
 we also want to show that fair queuing detection can be used to lower the queuing delay drastically while not impairing throughput. 
-\autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than two thirds while keeping throughput unchanged, when it is implemented on top on NewReno
+\autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than two thirds while keeping throughput unchanged, when it is implemented on top on NewReno. 
+However, Tonopah we want to emphasize that Tonopah is not a congestion control algorithm but just a mechanism to measure fair queuing which can be used with any congestion control algorithm.
 
@@
 
+As crosstraffic we use an iperf with the Newreno congestion control. We start it 4\,s so that it has time to saturate the link. 
+\autoref{table:no_fq_crosstraffic} and \autoref{table:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged.
+
 \subsection{Other variants of fair queuing}
-
-We also evaluated the performance of our fair queuing detection on a bottleneck managed by fq\_codel \cite{hoeiland-joergensen_flow_2018}. 
-We chose a default target queuing delay of 10\,ms following Apple's implementation\footnote{
-    \url{https://github.com/apple/darwin-xnu/blob/main/bsd/net/if.h\#L262}
-} because we argue that Apple probably spent a considerable amount of time fine-tuning their implementation and 
-came to the conclusion that 10\,ms work best as the default target delay.
 
@@
 
+We also evaluated the performance of our fair queuing detection on a bottleneck managed by fq\_codel \cite{hoeiland-joergensen_flow_2018}. 
+We chose a default target queuing delay of 10\,ms following Apple's implementation\footnote{
+    \url{https://github.com/apple/darwin-xnu/blob/main/bsd/net/if.h\#L262}
+} because we argue that Apple probably spent a considerable amount of time fine-tuning their implementation and 
+came to the conclusion that 10\,ms work best as the default target delay.
+
+\autoref{table:fq_codel} shows that Tonopah doesn't detect fair queuing as well when there's fq\_codel compared to when there is fq. Median detection accuracy is 95\%.
+
 \section{Discussion}

### 2022-06-15 sample/plot_bandwidth.py

@@ -37,11 +37,9 @@ for i, y in enumerate(ys):
 	plt.plot(x, y, alpha=0.67, label=labels[i])
 
 
-# plt.arrow(12.05, 10, 0, 12, color="black", edgecolor="black")
 plt.annotate("fair queuing", xy=(12.05, 23), xytext=(12.05, 10),
             arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
 
-# plt.arrow(12.05, 10, 0, 12, color="black", edgecolor="black")
 plt.annotate("fair queuing", xy=(13.4, 23), xytext=(13.4, 10),
             arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
 
@@ -50,5 +48,5 @@ plt.ylabel("Throughput (Mbit/s)")
 plt.ylim(bottom=0)
 plt.legend(loc="lower left")
 plt.tight_layout()
-plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
-plt.show()
+# plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.show()

### 2022-06-15 sample/plot_rtt.py

@@ -24,7 +24,7 @@ def add_plot(input_source):
     x = x / 1_000_000
     y = y / 1000.
     print("mean rtt:", np.mean(y))
-    plt.plot(x, y, alpha=0.67)
+    # plt.plot(x, y, alpha=0.67)
 
 if len(sys.argv) <= 1:
     input_source = fileinput.input()
@@ -40,4 +40,4 @@ else:
 plt.xlabel("Time (s)")
 plt.ylabel("Delay (ms)")
 plt.ylim(bottom=0)
-plt.show()
+# plt.show()

### 2022-06-16 picoquic/tex/paper.tex
Move tex

--- 2022-06-15 sample/tex/paper.tex
@@
 \centering
-\subfloat[\footnotesize \textbf{Sending} rate (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
+\subfloat[\footnotesize \textbf{Sending} rate at the sender (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
 \label{fig:sender}}\\
-\subfloat[\footnotesize \textbf{Receiving} rate (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
+\subfloat[\footnotesize \textbf{Receiving} rate at the receiver (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
 \label{fig:receiver}}
 \caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's \textbf{fair queuing}. 
-Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the flow that is sending more. 
+Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
 Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more (dominant subflow).
@@
 \centering
-\subfloat[\footnotesize \textbf{Sending} rate (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
+\subfloat[\footnotesize \textbf{Sending} rate at the sender (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
 \label{fig:sender_pfifo}}\\
-\subfloat[\footnotesize \textbf{Receiving} rate (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
+\subfloat[\footnotesize \textbf{Receiving} rate at the receiver (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
 \label{fig:receiver_pfifo}}
@@
 
-When different applications send packets on the internet, one application can send more than the other and thus unfairly take a larger share of bandwidth. 
+When different applications send packets over the internet, one application can send more than the other and thus unfairly take a larger share of bandwidth. 
 This can result in unfairness and bad user experience.
-Several different approaches have been proposed to address this \cite{brown_future_2020,ware_beyond_2019} : One is to make sure every network flow is ``well-behaved'' (also known as TCP friendly)
+Several different approaches have been proposed to address this \cite{brown_future_2020,ware_beyond_2019}: One is to make sure every network flow is ``well-behaved'' (also known as TCP friendly).
 Another one is to enforce fairness at switches and routers, called ``fair queuing'' or ``flow queuing'' \cite{nagle_packet_1985}.
 
-While fair queuing was proposed decades ago, it only gained popularity in the last couple of years because of implementations in the Linux kernel \cite{dumazet_pkt_sched_2013,hoeiland-joergensen_flow_2018}.
+While fair queuing was proposed decades ago, it only gained popularity in the last couple of years because of implementations in the Linux kernel \cite{dumazet_pkt_sched_2013,hoeiland-joergensen_flow_2018} and also in Apple's macOS.
 Applications can benefit from increasing deployment of fair queuing: It makes sure that not the most aggressive one wins. 
 It would be even better if applications could know if the connection they're sending on is managed by fair queuing. 
-Then they could be sure that the can use a common congestion control mechanism, while not bothering or being bothered by other network flows.
+Then they could be sure that the can use their preferred congestion control mechanism, while not bothering or being bothered by other network flows. 
 We proposed the first such approach in previous work \cite{bachl_detecting_2021} but our previous approach had some shortcomings upon which we improve in this paper.
 
-Our approach is a congestion control mechanism which also performs measurements to determine if there is fair queuing. 
-This approach of performing measurements in congestion control became popular in the last couple of years and was already followed by \cite{cardwell_bbr_2016,dong_pcc_2015,goyal_elasticity_2020,hayes_online_2020}.
-
-\section{Introduction}
-
-In our previous work we proposed a technique which determines the presence of fair queuing at flow startup \cite{bachl_detecting_2021}, which worked as follows: 
+In our previous work we proposed a technique which determines the presence of fair queuing at flow startup, which worked as follows: 
 If fair queuing is successfully detected at flow startup, a congestion control was used, which aimed to keep queuing delay low (delay-based congestion control). 
@@
 This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus is only used if fair queuing is detected. 
-If it is detected that there is no fair queuing, our approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other network flows,
+If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other network flows,
 but doesn't keep delay as low as our delay-based congestion control.
 
-While our previous approach had high detection accuracy (98\%), it also had some limitations:
+While our previous approach had high detection accuracy (98\%), it also had the following limitations:
 \begin{itemize}
@@
 
+Our new approach presented here is a mechanism which performs measurements to determine if there is fair queuing. Our mechanism can be added on top of any existing congestion control algorithm. 
+The approach of performing measurements in congestion control became popular in the last couple of years and was already followed by \cite{cardwell_bbr_2016,dong_pcc_2015,goyal_elasticity_2020,hayes_online_2020}.
+
 The core concept of our approach, which we call \textbf{\textit{Tonopah}}, is that a network flow is separated in \textbf{two subflows}. 
@@
 After a certain time interval the dominant subflow becomes the non-dominant one and vice versa. 
-When the total sending rate reaches the link capacity and fair queuing is present, \textbf{fair queuing} is going to \textbf{limit} the throughput of the \textbf{dominant subflow} (\autoref{fig:illustration_fq}).
+When the total sending rate reaches the link capacity and fair queuing is present, 
+\textbf{fair queuing} is going to \textbf{limit} the throughput of the \textbf{dominant subflow} (\autoref{fig:illustration_fq}). 
 This can be measured by observing the \textit{delivery rate}
@@
 Thus, in this paper, we define a new congestion signal: The presence of fair queuing. Other already known congestion signals are, for example: 
-packet loss; queue length, change of queue length, Explicit Congestion Notification \cite{mathis_relentless_2009,hutchison_revisiting_2011}.
+packet loss, queue length, change of queue length, Explicit Congestion Notification \cite{mathis_relentless_2009,hutchison_revisiting_2011}.
 
@@
 Another safeguard against measurement noise is that we only take a decision regarding fair queuing if the dominant flow actually really sent more data than the non-dominant one. 
-Specifically, if the dominant flow sends less than 57.5\%, we don't take a decision because the pacing apparently wasn't accurate enough. 
-This case only occurs when the sending rate is very low ($<$ 10\,Mbit/s) and thus pacing in inaccurate.
+Specifically, if the dominant flow sends less than 57.5\%, we don't take a decision because it means that probably we didn't have enough data to send. 
+This case only occurs when the sending rate is very low ($<$ 10\,Mbit/s).
 
@@
 We emulate the network using mininet and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing in \textit{fq} for fair queuing. 
+For evaluation we choose a network in which the sender is connected to the receiver via a switch (dumbbell topology).
+
+All the source code and evaluation code of this paper are freely available (refer to the very top of this paper).
 
 In the following evaluation we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
-If in the evaluation scenario there is fair queuing, and Tonopah detects fair queuing for 80\,s and fails to detect it for 10\,s we 
+If in the evaluation scenario there is fair queuing and Tonopah detects fair queuing for 80\,s and fails to detect it for 10\,s, we 
 state the accuracy as $\frac{80}{90} = 89\%$. 
@@
 \autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than two thirds while keeping throughput unchanged, when it is implemented on top on NewReno. 
-However, Tonopah we want to emphasize that Tonopah is not a congestion control algorithm but just a mechanism to measure fair queuing which can be used with any congestion control algorithm.
+However, we want to emphasize that Tonopah is not a congestion control algorithm but just a mechanism to measure fair queuing which can be used with any congestion control algorithm.
 
@@
 
-As crosstraffic we use an iperf with the Newreno congestion control. We start it 4\,s so that it has time to saturate the link. 
+To evaluate Tonopah's performance under crosstraffic, we use iperf with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
 \autoref{table:no_fq_crosstraffic} and \autoref{table:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged.
@@
 
+We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy. 
+We also showed that it can work with alternative versions of fair queuing such as fq\_codel although accuracy depends on the configuration of the queue manager, 
+the link capacity, base delay etc.
+Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cocoa} \cite{bachl_cocoa_2019} or \textit{cake} \cite{hoiland-jorgensen_piece_2018}.
+
+Another potentially fruitful line of research would be to integrate Tonopah into other congestion control algorithms such as BBR.
+
 \bibliographystyle{ieeetr}

### 2022-06-16 picoquic/tex/paper.tex
Proofread again

@@
 In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path. 
-When fair queuing is detected, a different congestion control is chosen, which results in reduced latency. 
+When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency. 
 Unlike an algorithm proposed in a previous paper of us, the approach presented here does not only detect the presence of fair queuing once at flow startup but it does so continuously. 
@@
 It would be even better if applications could know if the connection they're sending on is managed by fair queuing. 
-Then they could be sure that the can use their preferred congestion control mechanism, while not bothering or being bothered by other network flows. 
+Then they could be sure that they can use their preferred congestion control mechanism, while not bothering or being bothered by other network flows. 
 We proposed the first such approach in previous work \cite{bachl_detecting_2021} but our previous approach had some shortcomings upon which we improve in this paper.
@@
 similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}.
-This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus is only used if fair queuing is detected. 
+This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus it is only used if fair queuing is detected. 
 If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other network flows,
@@
 
-The core concept of our approach, which we call \textbf{\textit{Tonopah}}, is that a network flow is separated in \textbf{two subflows}. 
+The core concept of our approach, which we call \textbf{\textit{Tonopah}}, is that a network flow is separated into \textbf{two subflows}. 
 \textbf{Alternatingly}, one flow sends more (\textbf{dominant subflow}), while the other one sends less (\textbf{non-dominant subflow}). 
@@
 The dominant subflow and the non-dominant subflow alternate every $\max{(50, \min{(100, \textit{rtt})})}$ milliseconds. 
-This time period is also called an \textbf{interval}. To determine the presence of fair queuing, the data of the last four intervals are evaluated. 
+This time period is also called an \textbf{\textit{interval}}. To determine the presence of fair queuing, the data of the last four intervals are evaluated. 
 We look at more than one interval because this reduces the effect of measurement noise. We found four intervals to be a suitable number.
@@
 Specifically, if the dominant flow sends less than 57.5\%, we don't take a decision because it means that probably we didn't have enough data to send. 
-This case only occurs when the sending rate is very low ($<$ 10\,Mbit/s).
+This case only occurs when the underlying application has no data to send or when the sending rate is very low ($<$ 10\,Mbit/s).
 
@@
 
-Tonopah was implemented on top of Picquic in C. It was added on top of the NewReno congestion control algorithm but it could easily be added on top of other 
+Tonopah was implemented on top of Picoquic in C. It was added on top of the NewReno congestion control algorithm but it could easily be added on top of other 
 congestion control algorithms as well. By basing our implementation on Picoquic, Tonopah also supports Explicit Congestion Notification.
@@
 
-In the following evaluation we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
+In the following evaluations we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
 If in the evaluation scenario there is fair queuing and Tonopah detects fair queuing for 80\,s and fails to detect it for 10\,s, we 
@@
 \end{table}
+
+\subsection{Detection accuracy}
 

### 2022-06-17 tex/paper.tex
Adapt figures

--- 2022-06-16 picoquic/tex/paper.tex
@@
 \caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's \textbf{fair queuing}. 
-Around seconds 12 and 13.5, the sending rate reaches the maximum of the link -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
+Around seconds 82 and 83.5, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
 Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more (dominant subflow).
@@
 Even though the sender sends too much and a queue builds up, still the flow that sends more data also has a higher receiving rate. 
-This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing.}
+This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing. }
 \label{fig:illustration_no_fq}
@@
 state the accuracy as $\frac{80}{90} = 89\%$. 
+We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s. 
 \begin{table}
@@
 This means that Tonopah can deliver the same throughput while cause a lot less queuing delay. 
-The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value\protect\footnote{Using Welch's t-test} $< 10^{-5}$).}
+The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value\protect\footnote{Using Welch's t-test} $< 10^{-5}$). 
+The experiment setup is the same as in \autoref{table:fq}. }
 \label{table:throughput_and_delay}

### 2022-06-17 tex/paper.tex
Typo

@@
 \caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's \textbf{fair queuing}. 
-Around seconds 82 and 83.5, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
+Around second 82, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
 Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more (dominant subflow).

### 2022-06-17 sample/network_sim.py

@@ -68,11 +68,13 @@ bw_results = []
 delay_results = []
 results = []
 
-for delay in (10, 50, 100):
+# for delay in (10, 50, 100):
+for delay in (50,):
     bw_results.append([])
     delay_results.append([])
     results.append([])
-    for rate in (10, 50, 100):
+    # for rate in (10, 50, 100):
+    for rate in (50,):
         bw_results[-1].append([])
         delay_results[-1].append([])
         results[-1].append([])
@@ -257,9 +259,9 @@ for delay in (10, 50, 100):
                     info.append((ts, value))
             print("info", info)
             correct_duration = 0.0
-            if (unexpected_cnxid or len(info)-2 <= 0) and args.cc == "tonopah":
-                print("Got too few results...")
-                continue
+            # if (unexpected_cnxid or len(info)-2 <= 0) and args.cc == "tonopah":
+            #     print("Got too few results...")
+            #     continue
             if len(info)-2 > 0 and args.cc == "tonopah":    
                 for i in range(len(info)-2):
                     assert info[i+1][1] is not None
@@ -271,19 +273,19 @@ for delay in (10, 50, 100):
                 print("duration", duration, 'correct', correct_rate)
                 results[-1][-1].append(correct_rate)
 
-            bw_string = """tshark -n -r logs/client.pcap -q -z io,stat,0.01,"BYTES()udp.srcport == 4433 || udp.srcport == 4434","BYTES()udp.srcport == 4433","BYTES()udp.srcport == 4434","BYTES()ip.src==192.168.0.2" | grep '<>' | awk '{print $2","$6}' | sudo -u max python plot_bandwidth.py"""
-            output = subprocess.check_output(bw_string, shell=True).decode("utf-8")
-            parsed_bw = float(output.strip().split(' ')[-1])
-            print("bw", parsed_bw)
-            bw_results[-1][-1].append(parsed_bw)
-            delay_string = """cat logs/client.qlog | sudo -u max python plot_rtt.py"""
-            output = subprocess.check_output(delay_string, shell=True).decode("utf-8")
-            parsed_delay = float(output.strip().split(' ')[-1]) - delay
-            print("delay", parsed_delay)
-            delay_results[-1][-1].append(parsed_delay)
[…]

### 2022-06-17 sample/plot_bandwidth.py

@@ -37,16 +37,16 @@ for i, y in enumerate(ys):
 	plt.plot(x, y, alpha=0.67, label=labels[i])
 
 
-plt.annotate("fair queuing", xy=(12.05, 23), xytext=(12.05, 10),
-            arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
+# plt.annotate("fair queuing", xy=(82.05, 22), xytext=(82.05, 10),
+#             arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
 
-plt.annotate("fair queuing", xy=(13.4, 23), xytext=(13.4, 10),
-            arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
+# plt.annotate("fair queuing", xy=(83.5, 22), xytext=(83.5, 10),
+#             arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
 
 plt.xlabel("Time (s)")
 plt.ylabel("Throughput (Mbit/s)")
 plt.ylim(bottom=0)
-plt.legend(loc="lower left")
+plt.legend(loc="lower right")
 plt.tight_layout()
-# plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
-# plt.show()
+plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
+plt.show()

### 2022-06-17 sample/plot_rtt.py

@@ -24,7 +24,7 @@ def add_plot(input_source):
     x = x / 1_000_000
     y = y / 1000.
     print("mean rtt:", np.mean(y))
-    # plt.plot(x, y, alpha=0.67)
+    plt.plot(x, y, alpha=0.67)
 
 if len(sys.argv) <= 1:
     input_source = fileinput.input()
@@ -40,4 +40,4 @@ else:
 plt.xlabel("Time (s)")
 plt.ylabel("Delay (ms)")
 plt.ylim(bottom=0)
-# plt.show()
+plt.show()

### 2022-06-17 tex/paper.tex
Typo

@@
 Around second 82, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
-Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more (dominant subflow).
-This is the effect of fair queuing.}
+Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more. This is the effect of fair queuing.}
 \label{fig:illustration_fq}
@@
 \caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate in case there's \textbf{no fair queuing}. 
-Even though the sender sends too much and a queue builds up, still the flow that sends more data also has a higher receiving rate. 
-This is in contrast to \autoref{fig:illustration_fq}, where both flows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing. }
+Even though the sender sends too much and a queue builds up, still the subflow that sends more data also has a higher receiving rate. 
+This is in contrast to \autoref{fig:illustration_fq}, where both subflows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing. }
 \label{fig:illustration_no_fq}

### 2022-06-17 tex/paper.tex
refinement

@@
 If fair queuing is detected, the congestion control would not have to take cross traffic into account, which allows for more flexibility. 
-In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path. 
+In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path with very high accuracy. 
 When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency. 
@@
 When different applications send packets over the internet, one application can send more than the other and thus unfairly take a larger share of bandwidth.
+Also, a large queue can form if one flow sends more than it should.
 This can result in unfairness and bad user experience.
@@
 }
-of both subflows. If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}), which can be considered unfair. 
+of both subflows. If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}). 
 In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection. 
+This can be a potential advantage which a Tonopah-enabled congestion control can have over other congestion control algorithms such as BBR, which has been shown to be unfair to other flows sometimes \cite{ware_modeling_2019}.
 
@@
 Thus, in this paper, we define a new congestion signal: The presence of fair queuing. Other already known congestion signals are, for example: 
-packet loss, queue length, change of queue length, Explicit Congestion Notification \cite{mathis_relentless_2009,hutchison_revisiting_2011}.
+packet loss, queue length, change of queue length, Explicit Congestion Notification \cite{mathis_relentless_2009,hayes_revisiting_2011}.
 
@@
 
-We evaluated Tonopah on an Apple MacBook Air from 2011 with an Intel Core i5-2557M CPU at 1.70GHz. 
+We evaluated Tonopah on an Apple MacBook Air from 2011 with an Intel Core i5-2557M CPU at 1.7\,GHz. 
 By choosing old hardware for experimentation, we can show that our algorithm doesn't use excessive compute and can also run well on weak hardware.
 
-We emulate the network using mininet and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing in \textit{fq} for fair queuing. 
-For evaluation we choose a network in which the sender is connected to the receiver via a switch (dumbbell topology).
+We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing in \textit{fq} for fair queuing. 
+For evaluation we choose a network in which the sender is connected to the receiver via a switch.
 
@@
 We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s. 
+As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets. 
 \begin{table}
@@
 the link capacity, base delay etc.
-Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cocoa} \cite{bachl_cocoa_2019} or \textit{cake} \cite{hoiland-jorgensen_piece_2018}.
+Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cocoa} \cite{bachl_cocoa_2019}, \textit{LFQ} \cite{bachl_lfq_2020} or \textit{cake} \cite{hoiland-jorgensen_piece_2018}.
 

### 2022-06-19 sample/network_sim.py
Prettify code

@@ -39,7 +39,6 @@ class MyTopo(mininet.topo.Topo):
 
         # Add links
         self.addLink(leftHost, switch)
-        # self.addLink(leftHost, switch)
         self.addLink(switch, rightHost)
 
 topo = MyTopo()
@@ -68,13 +67,11 @@ bw_results = []
 delay_results = []
 results = []
 
-# for delay in (10, 50, 100):
-for delay in (50,):
+for delay in (10, 50, 100):
     bw_results.append([])
     delay_results.append([])
     results.append([])
-    # for rate in (10, 50, 100):
-    for rate in (50,):
+    for rate in (10, 50, 100):
         bw_results[-1].append([])
         delay_results[-1].append([])
         results[-1].append([])
@@ -87,9 +84,6 @@ for delay in (50,):
             print("delay", delay, "rate", rate, "rep_counter", rep_counter)
 
             opt = Opts()
-            # opt.qdisc = 'pfifo'
-            # opt.qdisc = 'fq_codel'
-            # opt.qdisc = 'fq'
             opt.qdisc = args.qdisc
             opt.delay = delay
             opt.rate = rate
@@ -100,10 +94,7 @@ for delay in (50,):
                 if with_delay:
                     print("bdp", bdp)
                 opt.buffer_size = None
-                # opt.buffer_size = int(1. * math.ceil(bdp))
-                # opt.buffer_size = 10
                 if opt.qdisc == 'pfifo' or opt.qdisc == 'fq':
-                    # opt.buffer_size = bdp
                     opt.buffer_size = max(100, bdp)
                 opt.interface = if_name
 
@@ -118,12 +109,8 @@ for delay in (50,):
                         if opt.buffer_size is not None: 
                             qdisc_string += f" flow_limit {int(opt.buffer_size)}"
                     elif opt.qdisc == 'fq_codel':
-                        # min_delay = (1500 * 8) / (rate * 10**6) * 100
-                        # fq_codel_delay = int(max(min_delay*1000, 10))
-                        # print('fq_codel_delay', fq_codel_delay)
                         fq_codel_delay = 10
                         qdisc_string = f"{opt.qdisc} target {fq_codel_delay}ms"
-                        # qdisc_string = f"{opt.qdisc}"
 
                 else:
                     qdisc_string = 'pfifo'
@@ -139,12 +126,9 @@ for delay in (50,):
                 return strings
 
[…]

### 2022-06-20 sample/network_sim.py
Update network_sim.py

@@ -79,7 +79,7 @@ for delay in (10, 50, 100):
         while True:
             subprocess.run('killall picoquic_sample'.split(' '))
             subprocess.run('killall iperf3'.split(' '))
-            if rep_counter >= 5:
+            if rep_counter >= 10:
                 break
             print("delay", delay, "rate", rate, "rep_counter", rep_counter)
 

### 2022-06-20 tex/paper.tex
Fix typo
Update paper.tex

@@
 If fair queuing is detected, the congestion control would not have to take cross traffic into account, which allows for more flexibility. 
-In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path with very high accuracy. 
+In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path, with very high accuracy. 
 When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency. 
@@
 & Link utilization & Queuing delay \\ \hline
-NewReno & 95.6\% & 10.5 ms \\ \hline
-Tonopah & 94.3\% & 33.6 ms \\ \hline
+NewReno & 95.6\% & 33.6 ms \\ \hline
+Tonopah & 94.3\% & 10.5 ms \\ \hline
 % p-value\protect\footnote{Using Welch's t-test} & 0.86 & $1.9 \times 10^{-6}$ \\ \hline
@@
 But the \textbf{queuing delay} of NewReno is \textbf{321\%} Tonopah's. 
-This means that Tonopah can deliver the same throughput while cause a lot less queuing delay. 
+This means that Tonopah can deliver the same throughput while causing a lot less queuing delay. 
 The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value\protect\footnote{Using Welch's t-test} $< 10^{-5}$). 

### 2022-06-21 tex/paper.tex
Spellchecking

@@
 \label{fig:receiver}}
-\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate in case there's \textbf{fair queuing}. 
+\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate of an example flow in case there's \textbf{fair queuing}. 
 Around second 82, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
-Thus, in the lower figure, the dominant subflow and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more. This is the effect of fair queuing.}
+Thus, in the lower figure, the dominant subflow (the one that is sending more) and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more.
+This is the effect of fair queuing.}
 \label{fig:illustration_fq}
@@
 \label{fig:receiver_pfifo}}
-\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate in case there's \textbf{no fair queuing}. 
+\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate of an example flow in case there's \textbf{no fair queuing}. 
 Even though the sender sends too much and a queue builds up, still the subflow that sends more data also has a higher receiving rate. 
@@
 
-In our previous work we proposed a technique which determines the presence of fair queuing at flow startup, which worked as follows: 
-If fair queuing is successfully detected at flow startup, a congestion control was used, which aimed to keep queuing delay low (delay-based congestion control). 
-While this delay-based congestion control achieved high throughput and low delay, it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,ha_cubic_2008},
-similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}.
+In our previous work we proposed a technique which determines the presence of fair queuing at flow startup, which works as follows: 
+If fair queuing is successfully detected at flow startup, a congestion control was used which aimed to keep queuing delay low (delay-based congestion control). 
+While this delay-based congestion control achieved high throughput and low delay, similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}, 
+it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,ha_cubic_2008}.
 This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus it is only used if fair queuing is detected. 
-If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other network flows,
+If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other, more aggressive network flows,
 but doesn't keep delay as low as our delay-based congestion control.
 
-While our previous approach had high detection accuracy (98\%), it also had the following limitations:
+While our previous approach had high detection accuracy (98\%), it also had some limitations:
 \begin{itemize}
-    \item It would \textbf{only} detect fair queuing at \textbf{flow startup}. 
+    \item It would \textbf{only} detect fair queuing only at \textbf{flow startup}. 
     But if the bottleneck link changes during a flow, it could be that the previous bottleneck had fair queuing while the new one doesn't. This wouldn't have been detected. 
@@
 }
-of both subflows. If there is no fair queuing, the subflow which sends more also achieves a higher throughput (\autoref{fig:illustration_no_fq}). 
+of both subflows. If there is no fair queuing, the subflow which sends more also achieves higher throughput (\autoref{fig:illustration_no_fq}). 
 In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection. 
@@
 The congestion window of the flow applies to both subflows together.
-This means that while always the dominant subflow sends more than the non-dominant one, in sum they always send the same amount of data, as allowed by the congestion window.
+This means that while always the dominant subflow sends more than the non-dominant one, in sum they never send more than what is allowed by the congestion window.
 
@@
 
-Another safeguard against measurement noise is that we only take a decision regarding fair queuing if the dominant flow actually really sent more data than the non-dominant one. 
-Specifically, if the dominant flow sends less than 57.5\%, we don't take a decision because it means that probably we didn't have enough data to send. 
+Another safeguard against measurement noise is that Tonopah only take a decision regarding fair queuing if the dominant flow actually really sent more data than the non-dominant one. 
+Specifically, if the dominant flow sends a fraction of throughput smaller than 0.575, Tonopah doesn't take a decision because it means that probably there wasn't enough data to send. 
 This case only occurs when the underlying application has no data to send or when the sending rate is very low ($<$ 10\,Mbit/s).
 
-Both subflows are controlled by the same congestion window. The differences in sending rates are implemented by using pacing. 
+Both subflows are controlled by the same congestion window. The differences in sending rates are achieved by using pacing. 
 The basis of the implementation of Tonopah is \textit{Picoquic}\footnote{\url{https://github.com/private-octopus/picoquic/tree/master/picoquic}}. 
@@
 
-Tonopah was implemented on top of Picoquic in C. It was added on top of the NewReno congestion control algorithm but it could easily be added on top of other 
+Tonopah was implemented on top of Picoquic in C. It was added on top of the NewReno congestion control algorithm but it could also be added on top of other 
 congestion control algorithms as well. By basing our implementation on Picoquic, Tonopah also supports Explicit Congestion Notification.
@@
 
-We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing in \textit{fq} for fair queuing. 
-For evaluation we choose a network in which the sender is connected to the receiver via a switch.
+We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} for fair queuing. 
+For evaluation we choose a simple network in which the sender is connected to the receiver via a switch.
 
-All the source code and evaluation code of this paper are freely available (refer to the very top of this paper).
+All the source code and evaluation code of this paper is freely available (refer to the very top of this paper).
 
 In the following evaluations we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
-If in the evaluation scenario there is fair queuing and Tonopah detects fair queuing for 80\,s and fails to detect it for 10\,s, we 
+If in the evaluation scenario there is fair queuing and Tonopah detects fair queuing correctly for 80\,s and fails to detect it for 10\,s, we 
 state the accuracy as $\frac{80}{90} = 89\%$. 
-We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s. 
+We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s in a grid-like fashion. 
 As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets. 
@@
 
-\autoref{table:fq_codel} shows that Tonopah doesn't detect fair queuing as well when there's fq\_codel compared to when there is fq. Median detection accuracy is 95\%.
+\autoref{table:fq_codel} shows that Tonopah doesn't detect fair queuing as well when there's fq\_codel compared to when there is fq. 
+However, median detection accuracy is still reasonably high at 95\%.
 
@@
 We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy. 
-We also showed that it can work with alternative versions of fair queuing such as fq\_codel although accuracy depends on the configuration of the queue manager, 
+We also showed that it can work with alternative versions of fair queuing such as fq\_codel although experiments showed that accuracy depends on the configuration of the queue manager, 
 the link capacity, base delay etc.

### 2022-06-21 tex/paper.tex
Update pdf

@@
 \usepackage{dblfloatfix}
+\usepackage{xurl}
+
 \renewcommand{\thefootnote}{\arabic{footnote}}
@@
 We chose a default target queuing delay of 10\,ms following Apple's implementation\footnote{
-    \url{https://github.com/apple/darwin-xnu/blob/main/bsd/net/if.h\#L262}
+    \url{https://github.com/apple/darwin-xnu/blob/2ff845c2e033bd0ff64b5b6aa6063a1f8f65aa32/bsd/net/if.h\#L262}
 } because we argue that Apple probably spent a considerable amount of time fine-tuning their implementation and 

### 2022-06-23 tex/paper.tex
Typo in results

@@
 & 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\%\\ \hline
+10ms & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
 50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-100ms & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
+100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% \\ \hline
 \end{tabularx}

### 2022-06-23 tex/paper.tex
Add reference

@@
 In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection. 
-This can be a potential advantage which a Tonopah-enabled congestion control can have over other congestion control algorithms such as BBR, which has been shown to be unfair to other flows sometimes \cite{ware_modeling_2019}.
+This can be a potential advantage which a Tonopah-enabled congestion control can have over other congestion control algorithms such as BBR, which has been shown to be unfair to other flows sometimes \cite{ware_modeling_2019,hock_experimental_2017}.
 

### 2022-07-02 tex/paper.tex
Update paper

@@
 As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets.
+
+\subsection{Detection accuracy}
+
 \begin{table}
@@
 \end{tabularx}
-\caption{Detection accuracy in case there's \textbf{no fair queuing}.  The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+\caption{Detection accuracy in case there's \textbf{no fair queuing}.  The overall median accuracy (true negative rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
 \label{table:no_fq}
@@
 \end{tabularx}
-\caption{Detection accuracy in case there is \textbf{fair queuing}.  The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+\caption{Detection accuracy in case there is \textbf{fair queuing}.  The overall median accuracy (true positive rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
 \label{table:fq}
@@
 
-\subsection{Detection accuracy}
-
-\autoref{table:no_fq} shows that if there is no fair queuing, Tonopah detects this correctly virtually always. 
-If there is fair queuing (\autoref{table:fq}), Tonopah detects this correctly very often but not in every case.
+\autoref{table:no_fq} shows that if there is no fair queuing, Tonopah detects this correctly virtually always (true negative rate $\geq$99\%). 
+If there is fair queuing (\autoref{table:fq}), Tonopah detects this correctly very often but not in every case (true positive rate in the high 90s).
 
@@
 \caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}.
-The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+The overall median accuracy (true negative rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
 \label{table:no_fq_crosstraffic}
@@
 \caption{Detection accuracy in case there is \textbf{fair queuing} under the presence of \textbf{cross-traffic}.
-The overall median accuracy is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
+The overall median accuracy (true positive rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
 \label{table:fq_crosstraffic}
@@
 
-\subsection{Other variants of fair queuing}
+\subsection{Other variants of fair queuing: fq\_codel}
 
@@
 \end{tabularx}
-\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing}.
-The overall median accuracy is 95\%, the first quartile is 83\% and the third quartile is 100\%.}
+\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing} with a target delay of 10\,ms.
+The overall median accuracy (true positive rate) is 95\%, the first quartile is 83\% and the third quartile is 100\%.}
 \label{table:fq_codel}
@@
 came to the conclusion that 10\,ms work best as the default target delay. 
+Also, Apple's implementation of fq\_codel seems to be the most widespread worldwide\footnote{\url{https://blog.cerowrt.org/post/state_of_fq_codel/}} with currently two billion Apple devices presumably supporting Apple's fq\_codel.
 
 \autoref{table:fq_codel} shows that Tonopah doesn't detect fair queuing as well when there's fq\_codel compared to when there is fq. 
-However, median detection accuracy is still reasonably high at 95\%.
+However, median detection accuracy (true positive rate) is still reasonably high at 95\%.
+
+\subsubsection{fq\_codel with a target delay of 5\,ms}
+
+\begin{table}
+\begin{tabularx}{\columnwidth}{| l | X | X | X |}
+\hline
+& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
+10ms & Median: 39\% \newline 1st quart.: 37\% \newline 3rd quart.: 80\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
+50ms & Median: 36\% \newline 1st quart.: 36\% \newline 3rd quart.: 36\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\% \\ \hline
+100ms & Median: 16\% \newline 1st quart.: 14\% \newline 3rd quart.: 17\% & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% & Median: 97\% \newline 1st quart.: 64\% \newline 3rd quart.: 99\% \\ \hline
+\end{tabularx}
+\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing} with a target delay of 5\,ms.
+The overall median accuracy (true positive rate) is 98\%, the first quartile is 37\% and the third quartile is 100\%.}
+\label{table:fq_codel_linux}
+\end{table}
+
+Although Apple's fq\_codel implementation with its 10\,ms default target delay seems to be the most widespread, fq\_codel is also the default queuing manager on multiple Linux distributions. 
+On Linux, fq\_codel has a default target delay of 5\,ms, half of Apple's value of 10.
+
+Since the target delay of 5\,ms is quite small, we had to slightly modify Tonopah: We fixed the interval of Tonopah to 50\,ms. 
+As a reminder, the interval is the time duration after which the dominant subflow becomes the non-dominant one and vice versa. 
+As a reminder, for the previous experiments, the interval was the current round trip time but at least 50\,ms and no more than 100\,ms.
+
+\autoref{table:fq_codel_linux} shows the performance of Tonopah with Linux's implementation of fq\_codel. Median detection accuracy (true positive rate) is 98\%. 
+A problem is that accuracy is pretty low for small bandwidths ($\leq$10\,Mbit/s) but otherwise fair queuing can be recognized quite accurately.
+
+Since we changed how Tonopah works to make it work better with Linux's implementation of fq\_codel by setting the interval to 50\,ms, we also have to make sure that this modification didn't lower the true negative rate (accuracy in case there is no fair queuing).
+Our experiments show that with the constant 50\,ms interval Tonopah even works slightly better, giving a true negative rate of 100\%, meaning it can always detect the absence of fair queuing correctly.
 
@@
 the link capacity, base delay etc.
-Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cocoa} \cite{bachl_cocoa_2019}, \textit{LFQ} \cite{bachl_lfq_2020} or \textit{cake} \cite{hoiland-jorgensen_piece_2018}.
+Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cake} \cite{hoiland-jorgensen_piece_2018}, \textit{PIE} \cite{pan_pie_2013}, \textit{cocoa} \cite{bachl_cocoa_2019} or \textit{LFQ} \cite{bachl_lfq_2020}.
 
 Another potentially fruitful line of research would be to integrate Tonopah into other congestion control algorithms such as BBR.
+
+\section*{Acknowledgements}
+
+We thank Sebastian Moeller, who suggested to also include experiments with fq\_codel with a target value of 5\,ms, which is the default on Linux.
 

### 2022-07-02 picoquic/tonopah.c

@@ -52,8 +52,8 @@ typedef struct st_picoquic_tonopah_interval_info_t {
 
 #define INTERVALS_REQUIRED 4
 uint64_t minimum_interval = 50000;
-uint64_t maximum_interval = 100000;
-// uint64_t maximum_interval = 50000;
+// uint64_t maximum_interval = 100000;
+uint64_t maximum_interval = 50000;
 
 picoquic_tonopah_interval_info_t* interval_list_first = NULL;
 picoquic_tonopah_interval_info_t* interval_list_last = NULL;

### 2022-07-02 sample/network_sim.py

@@ -79,7 +79,7 @@ for delay in (10, 50, 100):
         while True:
             subprocess.run('killall picoquic_sample'.split(' '))
             subprocess.run('killall iperf3'.split(' '))
-            if rep_counter >= 10:
+            if rep_counter >= 5:
                 break
             print("delay", delay, "rate", rate, "rep_counter", rep_counter)
 
@@ -109,8 +109,9 @@ for delay in (10, 50, 100):
                         if opt.buffer_size is not None: 
                             qdisc_string += f" flow_limit {int(opt.buffer_size)}"
                     elif opt.qdisc == 'fq_codel':
-                        fq_codel_delay = 10
-                        qdisc_string = f"{opt.qdisc} target {fq_codel_delay}ms"
+                        # fq_codel_delay = 10
+                        # qdisc_string = f"{opt.qdisc} target {fq_codel_delay}ms"
+                        qdisc_string = f"{opt.qdisc}"
 
                 else:
                     qdisc_string = 'pfifo'
@@ -218,7 +219,7 @@ for delay in (10, 50, 100):
                 if "Unexpected cnxid" in line:
                     uneected_cnxid = True
                     break
-                m = re.searxpch("Tonopah: (Ending|Recovery|FQ detected) at ([0-9]+)", line)
+                m = re.search("Tonopah: (Ending|Recovery|FQ detected) at ([0-9]+)", line)
                 if m:
                     if len(info) > 0 and info[-1][-1] == None:
                         break

### 2022-07-02 sample/plot_bandwidth.py

@@ -34,7 +34,7 @@ labels = ['total', 'subflow 1', 'subflow 2']
 for i, y in enumerate(ys):
 	print("sum:", np.sum(y) / 1_000_000 * 8)
 	y = y / interval_len / 1_000_000 * 8
-	plt.plot(x, y, alpha=0.67, label=labels[i])
+	# plt.plot(x, y, alpha=0.67, label=labels[i])
 
 
 # plt.annotate("fair queuing", xy=(82.05, 22), xytext=(82.05, 10),
@@ -43,10 +43,10 @@ for i, y in enumerate(ys):
 # plt.annotate("fair queuing", xy=(83.5, 22), xytext=(83.5, 10),
 #             arrowprops=dict(arrowstyle="->"), horizontalalignment='center')
 
-plt.xlabel("Time (s)")
-plt.ylabel("Throughput (Mbit/s)")
-plt.ylim(bottom=0)
-plt.legend(loc="lower right")
-plt.tight_layout()
-plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
-plt.show()
+# plt.xlabel("Time (s)")
+# plt.ylabel("Throughput (Mbit/s)")
+# plt.ylim(bottom=0)
+# plt.legend(loc="lower right")
+# plt.tight_layout()
+# plt.savefig("plots/bw.pdf", bbox_inches = 'tight', pad_inches = 0)
+# plt.show()

### 2022-09-20 tex/paper.tex
Clarify experiment description
Clarify experiment
Slightly adapt paper

@@
 When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency.
-Unlike an algorithm proposed in a previous paper of us, the approach presented here does not only detect the presence of fair queuing once at flow startup but it does so continuously. 
-% We show that by adding fair queuing detection to an existing congestion control algorithm, queuing delay can be reduced by more than two thirds. 
+In case fair queuing is detected, each flow can specify how much queuing delay it allows, meaning that each flow can choose its own tradeoff between throughput and latency.
 \end{abstract}
@@
 \centering
-\subfloat[\footnotesize \textbf{Sending} rate at the sender (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_fq"}.pdf}
-\label{fig:sender}}\\
-\subfloat[\footnotesize \textbf{Receiving} rate at the receiver (fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_fq"}.pdf}
-\label{fig:receiver}}
-\caption{Figure \ref{fig:sender} shows the sending rate, \ref{fig:receiver} the receiving rate of an example flow in case there's \textbf{fair queuing}. 
-Around second 82, the total sending rate is exceeding the link capacity -- 50 Mbit/s -- and fair queuing starts to limit the throughput of the subflow that is sending more. 
-Thus, in the lower figure, the dominant subflow (the one that is sending more) and the non-dominant subflow achieve approx.~the same receiving rate even though the dominant subflow sends more.
-This is the effect of fair queuing.}
+\includegraphics[width=\columnwidth]{{"../sample/plots/concept_fq"}.pdf}
+\caption{In case there is \textbf{fair queuing}, when the bottleneck link is congested, subflow 1, which sends more, sees increasing queuing delay.
+Subflow 2 doesn't see increased queuing delay as it sends less than its fair share and thus its packets don't have to wait at the bottleneck link.}
 \label{fig:illustration_fq}
@@
 \centering
-\subfloat[\footnotesize \textbf{Sending} rate at the sender (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_server_pfifo"}.pdf}
-\label{fig:sender_pfifo}}\\
-\subfloat[\footnotesize \textbf{Receiving} rate at the receiver (no fair queuing)]{\includegraphics[width=\columnwidth]{{"figures/bw_client_pfifo"}.pdf}
-\label{fig:receiver_pfifo}}
-\caption{Figure \ref{fig:sender_pfifo} shows the sending rate, \ref{fig:receiver_pfifo} the receiving rate of an example flow in case there's \textbf{no fair queuing}. 
-Even though the sender sends too much and a queue builds up, still the subflow that sends more data also has a higher receiving rate. 
-This is in contrast to \autoref{fig:illustration_fq}, where both subflows have the same receiving rate once there is congestion at the bottleneck, thanks to fair queuing. }
+\includegraphics[width=\columnwidth]{{"../sample/plots/concept_fcfs"}.pdf}
+\caption{In case there is \textbf{no fair queuing}, when the bottleneck link is congested, a queue builds up. Since both subflows share the same queue, they both see increased queuing delay.}
 \label{fig:illustration_no_fq}
@@
 We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s in a grid-like fashion. 
+For each combination of delay and link capacity, multiple experiment runs are performed and the median, the quartiles or the average of the results are shown. 
 As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets.

### 2022-09-21 tex/paper.tex
Continue refining

@@
     \item can be \textbf{transparently used} on top of any congestion control algorithm. 
+    \item lets each application \textbf{choose how much delay} it allows in case fair queuing is detected. Allowing a higher delay can result in higher throughput. 
 \end{itemize}
@@
 The core concept of our approach, which we call \textbf{\textit{Tonopah}}, is that a network flow is separated into \textbf{two subflows}. 
-\textbf{Alternatingly}, one flow sends more (\textbf{dominant subflow}), while the other one sends less (\textbf{non-dominant subflow}). 
-After a certain time interval the dominant subflow becomes the non-dominant one and vice versa. 
+One flow constantly sends more data (\textbf{dominant subflow}), while the other one sends less (\textbf{non-dominant subflow}). 
 When the total sending rate reaches the link capacity and fair queuing is present, 
 \textbf{fair queuing} is going to \textbf{limit} the throughput of the \textbf{dominant subflow} (\autoref{fig:illustration_fq}). 
-This can be measured by observing the \textit{delivery rate}
-\footnote{The delivery rate (a concept popularized by BBR \cite{cardwell_bbr_2016}) can be measured by observing the rate at which acknowledgement packets are sent back to the sender by the receiver.
-If the sender sends more than the link capacity, the delivery rate is never going to exceed the link capacity since the receiver cannot receive packets faster than the link capacity allows. 
-}
-of both subflows. If there is no fair queuing, the subflow which sends more also achieves higher throughput (\autoref{fig:illustration_no_fq}). 
+Since its throughput is limited, a queue is going to build up and delay is going to increase.
+If the delay only increases for the dominant subflow but not for the non-dominant one, this means that there's fair queuing.
+
+On the other side, if there is no fair queuing, when the sending rate exceeds the capacity of the bottleneck link, a queuing is going to build up and queuing delay is going to rise for \textbf{both} the dominant and the non-dominant subflow (\autoref{fig:illustration_no_fq}).
 In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection. 
 This can be a potential advantage which a Tonopah-enabled congestion control can have over other congestion control algorithms such as BBR, which has been shown to be unfair to other flows sometimes \cite{ware_modeling_2019,hock_experimental_2017}.
-
-By alternating which subflow is dominant and which one is non-dominant, we prevent the build-up of a standing queue at the bottleneck: 
-When the dominant subflow builds up a queue at the bottleneck link, it is going to be the non-dominant subflow in the next time interval and then it is going to send less and the queue is going to be reduced.
-
-The congestion window of the flow applies to both subflows together.
-This means that while always the dominant subflow sends more than the non-dominant one, in sum they never send more than what is allowed by the congestion window.
 
@@
 An interesting aspect of fair queuing detection is that the presence of fair queuing can be interpreted as a \textit{congestion signal}: 
-Only if there is congestion, fair queuing can be detected. If the link is underutilized, flows don't ``fight'' for bandwidh and thus fairness doesn't need to be enforced at the bottleneck link. 
+Only if there is congestion, fair queuing can be detected. If the link is underutilized, flows don't ``fight'' for bandwidth and thus fairness doesn't need to be enforced at the bottleneck link. 
 Thus, in this paper, we define a new congestion signal: The presence of fair queuing. Other already known congestion signals are, for example: 
@@
 
-The dominant subflow receives $\frac{6}{10}$ of the total allowed sending rate, while the non-dominant one receives $\frac{4}{10}$. 
-The dominant subflow and the non-dominant subflow alternate every $\max{(50, \min{(100, \textit{rtt})})}$ milliseconds. 
-This time period is also called an \textbf{\textit{interval}}. To determine the presence of fair queuing, the data of the last four intervals are evaluated. 
-We look at more than one interval because this reduces the effect of measurement noise. We found four intervals to be a suitable number.
+The dominant subflow receives $\frac{2}{3}$ of the total allowed sending rate, while the non-dominant one receives the remaining $\frac{1}{3}$.
 
-We determine that fair queuing is present if the receiving rate was more fair than not fair. 
-It would be completely fair if both flows get $\frac{1}{2}$ of the total throughput. It would be completely unfair if the dominant subflow got $\frac{6}{10}$ and the non-dominant one $\frac{4}{10}$. 
-If the dominant one got a fraction of 0.549 of the total throughput, it would be just a little bit more fair than unfair. In this case we would say that there is fair queuing. 
-If, on the other side, the dominant subflow got a fraction of 0.551 of the total throughput, we would say that there is no fair queuing.
-
-Another safeguard against measurement noise is that Tonopah only take a decision regarding fair queuing if the dominant flow actually really sent more data than the non-dominant one. 
-Specifically, if the dominant flow sends a fraction of throughput smaller than 0.575, Tonopah doesn't take a decision because it means that probably there wasn't enough data to send. 
-This case only occurs when the underlying application has no data to send or when the sending rate is very low ($<$ 10\,Mbit/s).
+After each round-trip time, Tonopah checks if the queuing delay of dominant subflow was on average larger than the one of the non-dominant subflow.
+As the threshold we use 5\,ms: If the average queuing delay in the last round-trip time was 5\,ms higher for the dominant subflow compared to the non-dominant one, fair queuing is detected.
+However, it is worth noting that each application can choose this threshold individually.
+For example, a cloud gaming application might choose a threshold of 5\,ms and thus prioritize low delay over high throughput.
+However, a video chat application might want to emphasize high bandwidth more and thus choose a threshold of 20\,ms to get maximum throughput.
 

### 2022-09-22 tex/paper.tex
Continue refining

@@
 For each combination of delay and link capacity, multiple experiment runs are performed and the median, the quartiles or the average of the results are shown. 
-As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets.
+As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets when no fair queuing (pfifo) is used.
+When using fq for fair queuing, we set the buffer size for each flow using the same rule.
+When using fair queuing with fq\_codel, we use it in its default configuration.
 
@@
 
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there's \textbf{no fair queuing}.  The overall median accuracy (true negative rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
-\label{table:no_fq}
-\end{table}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_all"}.pdf}
+\caption{Detection accuracy in case there's \textbf{no fair queuing} (pfifo). The overall accuracy is 99.4\%.}
+\label{fig:no_fq}
+\end{figure}
 
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there is \textbf{fair queuing}.  The overall median accuracy (true positive rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
-\label{table:fq}
-\end{table}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah_all"}.pdf}
+\caption{Detection accuracy in case there is \textbf{fair queuing} (fq\_codel). The overall accuracy is 95.8\%.}
+\label{fig:fq_codel}
+\end{figure}
 
-\autoref{table:no_fq} shows that if there is no fair queuing, Tonopah detects this correctly virtually always (true negative rate $\geq$99\%). 
-If there is fair queuing (\autoref{table:fq}), Tonopah detects this correctly very often but not in every case (true positive rate in the high 90s).
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_new_tonopah_all"}.pdf}
+\caption{Detection accuracy in case there is \textbf{fair queuing} (fq). The overall accuracy is 95.4\%.}
+\label{fig:fq}
+\end{figure}
+
+\autoref{fig:no_fq} shows that if there is no fair queuing, Tonopah detects this correctly virtually always. 
+If there is fair queuing with fq\_codel (\autoref{fig:fq_codel}) or fq (\autoref{fig:fq}), Tonopah detects this correctly very often but not in every case.
 
@@
 & Link utilization & Queuing delay \\ \hline
-NewReno & 95.6\% & 33.6 ms \\ \hline
-Tonopah & 94.3\% & 10.5 ms \\ \hline
-% p-value\protect\footnote{Using Welch's t-test} & 0.86 & $1.9 \times 10^{-6}$ \\ \hline
-% $\frac{\text{NewReno}}{\text{Tonopah}}$ & 103\% & 321\% \\ \hline
+NewReno & 95.6\% & 36.5 ms \\ \hline
+Tonopah & 91.1\% & 3.3 ms \\ \hline
 \end{tabular}
-\caption{NewReno's link utilization is 103\% Tonopah's. 
-But the \textbf{queuing delay} of NewReno is \textbf{321\%} Tonopah's. 
-This means that Tonopah can deliver the same throughput while causing a lot less queuing delay. 
-The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value\protect\footnote{Using Welch's t-test} $< 10^{-5}$). 
-The experiment setup is the same as in \autoref{table:fq}. }
+\caption{NewReno's link utilization is 105\% Tonopah's. 
+But the \textbf{queuing delay} of NewReno is \textbf{1111\%} Tonopah's. 
+This means that Tonopah can deliver the same or slightly less throughput while causing a lot less queuing delay. 
+The differences in \textbf{queuing delay} are \textbf{highly significant} (p-value (Welch's t-test) $< 10^{-21}$). 
+The experiment setup is the same as in \autoref{fig:fq}. }
 \label{table:throughput_and_delay}
@@
 
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}.
-The overall median accuracy (true negative rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
-\label{table:no_fq_crosstraffic}
-\end{table}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf_all"}.pdf}
+\caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}. The overall accuracy is 99.9\%.}
+\label{fig:no_fq_crosstraffic}
+\end{figure}
 
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-100ms & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there is \textbf{fair queuing} under the presence of \textbf{cross-traffic}.
-The overall median accuracy (true positive rate) is 100\%, the first quartile is 100\% and the third quartile is 100\%.}
-\label{table:fq_crosstraffic}
-\end{table}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf_all"}.pdf}
+\caption{Detection accuracy in case there is \textbf{fair queuing} (fq) under the presence of \textbf{cross-traffic}. The overall accuracy is 98.1\%.}
+\label{fig:fq_crosstraffic}
+\end{figure}
 
 To evaluate Tonopah's performance under crosstraffic, we use iperf with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
-\autoref{table:no_fq_crosstraffic} and \autoref{table:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged.
-
-\subsection{Other variants of fair queuing: fq\_codel}
-
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 85\% \newline 1st quart.: 84\% \newline 3rd quart.: 85\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 91\% \newline 1st quart.: 88\% \newline 3rd quart.: 92\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 98\% \newline 1st quart.: 97\% \newline 3rd quart.: 99\% \\ \hline
-100ms & Median: 9\% \newline 1st quart.: 7\% \newline 3rd quart.: 13\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\% & Median: 64\% \newline 1st quart.: 0\% \newline 3rd quart.: 75\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing} with a target delay of 10\,ms.
-The overall median accuracy (true positive rate) is 95\%, the first quartile is 83\% and the third quartile is 100\%.}
-\label{table:fq_codel}
-\end{table}
-
-We also evaluated the performance of our fair queuing detection on a bottleneck managed by fq\_codel \cite{hoeiland-joergensen_flow_2018}. 
-We chose a default target queuing delay of 10\,ms following Apple's implementation\footnote{
-    \url{https://github.com/apple/darwin-xnu/blob/2ff845c2e033bd0ff64b5b6aa6063a1f8f65aa32/bsd/net/if.h\#L262}
-} because we argue that Apple probably spent a considerable amount of time fine-tuning their implementation and 
-came to the conclusion that 10\,ms work best as the default target delay. 
-Also, Apple's implementation of fq\_codel seems to be the most widespread worldwide\footnote{\url{https://blog.cerowrt.org/post/state_of_fq_codel/}} with currently two billion Apple devices presumably supporting Apple's fq\_codel.
-
-\autoref{table:fq_codel} shows that Tonopah doesn't detect fair queuing as well when there's fq\_codel compared to when there is fq. 
-However, median detection accuracy (true positive rate) is still reasonably high at 95\%.
-
-\subsubsection{fq\_codel with a target delay of 5\,ms}
-
-\begin{table}
-\begin{tabularx}{\columnwidth}{| l | X | X | X |}
-\hline
-& 10 Mbit/s & 50 Mbit/s & 100 Mbit/s \\ \hline
-10ms & Median: 39\% \newline 1st quart.: 37\% \newline 3rd quart.: 80\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 100\% \newline 1st quart.: 100\% \newline 3rd quart.: 100\%\\ \hline
-50ms & Median: 36\% \newline 1st quart.: 36\% \newline 3rd quart.: 36\% & Median: 100\% \newline 1st quart.: 99\% \newline 3rd quart.: 100\% & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 99\% \\ \hline
-100ms & Median: 16\% \newline 1st quart.: 14\% \newline 3rd quart.: 17\% & Median: 99\% \newline 1st quart.: 98\% \newline 3rd quart.: 100\% & Median: 97\% \newline 1st quart.: 64\% \newline 3rd quart.: 99\% \\ \hline
-\end{tabularx}
-\caption{Detection accuracy in case there is the \textbf{\textit{fq\_codel}} variant of \textbf{fair queuing} with a target delay of 5\,ms.
-The overall median accuracy (true positive rate) is 98\%, the first quartile is 37\% and the third quartile is 100\%.}
-\label{table:fq_codel_linux}
-\end{table}
-
-Although Apple's fq\_codel implementation with its 10\,ms default target delay seems to be the most widespread, fq\_codel is also the default queuing manager on multiple Linux distributions. 
-On Linux, fq\_codel has a default target delay of 5\,ms, half of Apple's value of 10.
-
-Since the target delay of 5\,ms is quite small, we had to slightly modify Tonopah: We fixed the interval of Tonopah to 50\,ms. 
-As a reminder, the interval is the time duration after which the dominant subflow becomes the non-dominant one and vice versa. 
-As a reminder, for the previous experiments, the interval was the current round trip time but at least 50\,ms and no more than 100\,ms.
-
-\autoref{table:fq_codel_linux} shows the performance of Tonopah with Linux's implementation of fq\_codel. Median detection accuracy (true positive rate) is 98\%. 
-A problem is that accuracy is pretty low for small bandwidths ($\leq$10\,Mbit/s) but otherwise fair queuing can be recognized quite accurately.
-
-Since we changed how Tonopah works to make it work better with Linux's implementation of fq\_codel by setting the interval to 50\,ms, we also have to make sure that this modification didn't lower the true negative rate (accuracy in case there is no fair queuing).
-Our experiments show that with the constant 50\,ms interval Tonopah even works slightly better, giving a true negative rate of 100\%, meaning it can always detect the absence of fair queuing correctly.
+\autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged or even slightly better.
+We tried to use fq\_codel as in \autoref{fig:fq} but we encountered problems when using fq\_codel in combination with iperf3.
+Frequently, iperf3 would lose all packets, have a timeout after a couple of seconds and stop.
+Since we could not resolve this issue, we used fq instead of fq\_codel, for which this issue didn't occur.
+For fq, we set the buffer to the same size as for pfifo (1 bandwidth delay product but at least 100 packets).
 
@@
 
-We thank Sebastian Moeller, who suggested to also include experiments with fq\_codel with a target value of 5\,ms, which is the default on Linux.
+We thank Sebastian Moeller, who suggested to add CDF plots and to use fq\_codel with a target value of 5\,ms, which is the default on Linux.
 

### 2022-09-25 tex/paper.tex
Add figures

@@
 \centering 
+\subfloat[\footnotesize A host sending data to a receiver. There is a bottleneck link with a switch which applies \textbf{fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/fq"}.pdf}
+\label{fig:fq_overview}}\\
+\subfloat[\footnotesize Queuing delay in case of \textbf{fair queuing}.]{
 \includegraphics[width=\columnwidth]{{"../sample/plots/concept_fq"}.pdf}
+\label{fig:fq_detail}}
 \caption{In case there is \textbf{fair queuing}, when the bottleneck link is congested, subflow 1, which sends more, sees increasing queuing delay.
@@
 \centering
+\subfloat[\footnotesize A host sending data to a receiver. There is a bottleneck link with a switch which applies \textbf{no fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/no_fq"}.pdf}
+\label{fig:no_fq_overview}}\\
+\subfloat[\footnotesize Queuing delay in case there is \textbf{no fair queuing}.]{
 \includegraphics[width=\columnwidth]{{"../sample/plots/concept_fcfs"}.pdf}
+\label{fig:no_fq_detail}}
 \caption{In case there is \textbf{no fair queuing}, when the bottleneck link is congested, a queue builds up. Since both subflows share the same queue, they both see increased queuing delay.}

### 2022-09-28 tex/paper.tex
Proofread

@@
 When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency.
-In case fair queuing is detected, each flow can specify how much queuing delay it allows, meaning that each flow can choose its own tradeoff between throughput and latency.
+Also, each flow can then specify how much queuing delay it allows, meaning that it can choose its own tradeoff between throughput and latency.
 \end{abstract}
@@
 This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus it is only used if fair queuing is detected. 
-If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (specifically PCC \cite{dong_pcc_2015}), which can compete better with other, more aggressive network flows,
+If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (PCC \cite{dong_pcc_2015}), which can compete better with other, more aggressive network flows,
 but doesn't keep delay as low as our delay-based congestion control.
@@
 
-We would thus like to have an approach which 
+We would like to have an approach which 
 \begin{itemize}
@@
     \item can be \textbf{transparently used} on top of any congestion control algorithm. 
-    \item lets each application \textbf{choose how much delay} it allows in case fair queuing is detected. Allowing a higher delay can result in higher throughput. 
+    \item lets each application \textbf{choose how much delay} it allows in case fair queuing is detected. Allowing a higher queuing delay can result in higher throughput. 
 \end{itemize}
@@
 
-On the other side, if there is no fair queuing, when the sending rate exceeds the capacity of the bottleneck link, a queuing is going to build up and queuing delay is going to rise for \textbf{both} the dominant and the non-dominant subflow (\autoref{fig:illustration_no_fq}).
+On the other side, if there is no fair queuing, when the sending rate exceeds the capacity of the bottleneck link, a queue is going to build up and queuing delay is going to rise for \textbf{both} the dominant and the non-dominant subflow (\autoref{fig:illustration_no_fq}).
 In this case, Tonopah doesn't behave different from a network flow which doesn't have fair queuing detection. 
@@
 The basis of the implementation of Tonopah is \textit{Picoquic}\footnote{\url{https://github.com/private-octopus/picoquic/tree/master/picoquic}}. 
-Picoquic includes support for Multipath QUIC and pacing, which we need for the implementation of Tonopah. Also, it implements congestion control algorithms. 
+Picoquic includes support for Multipath QUIC and pacing, which we need for the implementation of Tonopah. Also, Picoquic implements several congestion control algorithms. 
 We base Tonopah on the NewReno implementation of Picquic. When fair queuing is detected, the congestion window is reduced by $\frac{1}{8}$. 
@@
 
-We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} for fair queuing. 
-For evaluation we choose a simple network in which the sender is connected to the receiver via a switch.
+We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} \textit{fq\_codel} for fair queuing. 
+For the evaluation we choose a simple network in which the sender is connected to the receiver via a switch.
 
@@
 
-To evaluate Tonopah's performance under crosstraffic, we use iperf with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
+To evaluate Tonopah's performance under crosstraffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
 \autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged or even slightly better.
@@
 Since we could not resolve this issue, we used fq instead of fq\_codel, for which this issue didn't occur.
-For fq, we set the buffer to the same size as for pfifo (1 bandwidth delay product but at least 100 packets).
 

### 2022-09-29 tex/paper.tex
Add ecdfs for all bandwidths/delays combined

@@
 
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah"}.pdf}
+\label{fig:ecdf_pfifo}}
+\subfloat[\footnotesize fq\_codel]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah"}.pdf}
+\label{fig:ecdf_fq_codel}}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah"}.pdf}
+\label{fig:ecdf_fq}}
+\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing.}
+\label{fig:ecdf}
+\end{figure}
+
 \begin{figure}
@@
 If there is fair queuing with fq\_codel (\autoref{fig:fq_codel}) or fq (\autoref{fig:fq}), Tonopah detects this correctly very often but not in every case.
+\autoref{fig:ecdf} shows the empirical cumulative distribution functions.
 
@@
 
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_pfifo}}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_fq}}
+\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing when there is cross-traffic.}
+\label{fig:ecdf_crosstraffic}
+\end{figure}
+
 \begin{figure}
@@
 
-To evaluate Tonopah's performance under crosstraffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
+To evaluate Tonopah's performance under cross-traffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
 \autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged or even slightly better.
@@
 Since we could not resolve this issue, we used fq instead of fq\_codel, for which this issue didn't occur.
+
+\autoref{fig:ecdf_crosstraffic} shows the empirical cumulative distribution functions when there is cross-traffic.
 

### 2022-09-30 tex/paper.tex
Proofread

@@
 \begin{itemize}
-    \item It would \textbf{only} detect fair queuing only at \textbf{flow startup}. 
+    \item It would \textbf{only} detect fair queuing at \textbf{flow startup}. 
     But if the bottleneck link changes during a flow, it could be that the previous bottleneck had fair queuing while the new one doesn't. This wouldn't have been detected. 
@@
 
-We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} \textit{fq\_codel} for fair queuing. 
+We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} and \textit{fq\_codel} for fair queuing. 
 For the evaluation we choose a simple network in which the sender is connected to the receiver via a switch.
@@
 We evaluate the performance of Tonopah over a range of delays from 10 to 100\,ms and link capacity from 10 to 100\,Mbit/s in a grid-like fashion. 
-For each combination of delay and link capacity, multiple experiment runs are performed and the median, the quartiles or the average of the results are shown. 
+For each combination of delay and link capacity, 10 experiment runs are performed and the empirical cumulative distribution functions and the average detection accuracy are shown.
+
 As the buffer size at the bottleneck we set the bandwidth delay product but at least 100 packets when no fair queuing (pfifo) is used.
@@
 \subsection{Detection accuracy}
-
-\begin{figure}[h]
-\centering
-\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah"}.pdf}
-\label{fig:ecdf_pfifo}}
-\subfloat[\footnotesize fq\_codel]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah"}.pdf}
-\label{fig:ecdf_fq_codel}}
-\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah"}.pdf}
-\label{fig:ecdf_fq}}
-\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing.}
-\label{fig:ecdf}
-\end{figure}
 
@@
 \label{fig:fq}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah"}.pdf}
+\label{fig:ecdf_pfifo}}
+\subfloat[\footnotesize fq\_codel]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah"}.pdf}
+\label{fig:ecdf_fq_codel}}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah"}.pdf}
+\label{fig:ecdf_fq}}
+\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing.}
+\label{fig:ecdf}
 \end{figure}
@@
 While the goal of this paper is to implement a mechanism to detect fair queuing in an online fashion, 
-we also want to show that fair queuing detection can be used to lower the queuing delay drastically while not impairing throughput. 
-\autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than two thirds while keeping throughput unchanged, when it is implemented on top on NewReno. 
+we also show that fair queuing detection can be used to lower the queuing delay drastically while not impairing throughput. 
+\autoref{table:throughput_and_delay} shows that Tonopah can lower queuing delay by more than $\frac{9}{10}$ while only achieving slightly lower throughput, when it is implemented on top on NewReno. 
 However, we want to emphasize that Tonopah is not a congestion control algorithm but just a mechanism to measure fair queuing which can be used with any congestion control algorithm.
@@
 \subsection{Cross-traffic}
-
-\begin{figure}[h]
-\centering
-\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf"}.pdf}
-\label{fig:ecdf_crosstraffic_pfifo}}
-\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf"}.pdf}
-\label{fig:ecdf_crosstraffic_fq}}
-\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing when there is cross-traffic.}
-\label{fig:ecdf_crosstraffic}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_pfifo}}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_fq}}
+\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing when there is \textbf{cross-traffic}.}
+\label{fig:ecdf_crosstraffic}
+\end{figure}
+
 To evaluate Tonopah's performance under cross-traffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
-\autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic Tonpah's detection accuracy is virtually unchanged or even slightly better.
+\autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic, Tonpah's detection accuracy is virtually unchanged or even slightly better.
 We tried to use fq\_codel as in \autoref{fig:fq} but we encountered problems when using fq\_codel in combination with iperf3.
@@
 
-We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy. 
-We also showed that it can work with alternative versions of fair queuing such as fq\_codel although experiments showed that accuracy depends on the configuration of the queue manager, 
+We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy for both fq and fq\_codel, although experiments showed that accuracy depends on the configuration of the queue manager,
 the link capacity, base delay etc.

### 2022-10-02 tex/paper.tex
Proofread

@@
 \centering 
-\subfloat[\footnotesize A host sending data to a receiver. There is a bottleneck link with a switch which applies \textbf{fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/fq"}.pdf}
+\subfloat[\footnotesize A host sending data to a receiver via two concurrent flows.
+The blue flow sends two times more data.
+There is a bottleneck link with a switch which applies \textbf{fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/fq"}.pdf}
 \label{fig:fq_overview}}\\
@@
 \label{fig:fq_detail}}
-\caption{In case there is \textbf{fair queuing}, when the bottleneck link is congested, subflow 1, which sends more, sees increasing queuing delay.
+\caption{In case there is \textbf{fair queuing}, when the bottleneck link is congested, subflow 1, which sends more, sees increased queuing delay.
 Subflow 2 doesn't see increased queuing delay as it sends less than its fair share and thus its packets don't have to wait at the bottleneck link.}
@@
 \centering
-\subfloat[\footnotesize A host sending data to a receiver. There is a bottleneck link with a switch which applies \textbf{no fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/no_fq"}.pdf}
+\subfloat[\footnotesize A host sending data to a receiver via two concurrent flows.
+The blue flow sends two times more data.
+There is a bottleneck link with a switch which applies \textbf{no fair queuing}.]{\includegraphics[width=\columnwidth]{{"figures/no_fq"}.pdf}
 \label{fig:no_fq_overview}}\\
@@
 While fair queuing was proposed decades ago, it only gained popularity in the last couple of years because of implementations in the Linux kernel \cite{dumazet_pkt_sched_2013,hoeiland-joergensen_flow_2018} and also in Apple's macOS.
-Applications can benefit from increasing deployment of fair queuing: It makes sure that not the most aggressive one wins. 
+Applications can benefit from increased deployment of fair queuing: It makes sure that not the most aggressive one wins. 
 It would be even better if applications could know if the connection they're sending on is managed by fair queuing. 
@@
 In our previous work we proposed a technique which determines the presence of fair queuing at flow startup, which works as follows: 
-If fair queuing is successfully detected at flow startup, a congestion control was used which aimed to keep queuing delay low (delay-based congestion control). 
-While this delay-based congestion control achieved high throughput and low delay, similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}, 
-it was vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,ha_cubic_2008}.
-This means that our delay-based congestion control performed well but only when it wouldn't have to compete with other flows. Thus it is only used if fair queuing is detected. 
+If fair queuing is successfully detected at flow startup, a congestion control is used which aims to keep queuing delay low (delay-based congestion control). 
+While this delay-based congestion control achieves high throughput and low delay, similar to the Vegas congestion control algorithm \cite{brakmo_tcp_1995}, 
+it is vulnerable to be outcompeted by other network flows sending more aggressively, such as \cite{cardwell_bbr_2016,dong_pcc_2015,ha_cubic_2008}.
+This means that our delay-based congestion control performs well but only when it doesn't have to compete with other flows. Thus it is only used if fair queuing is detected. 
 If the absence of fair queuing is detected, our previous approach uses a more aggressive congestion control (PCC \cite{dong_pcc_2015}), which can compete better with other, more aggressive network flows,
@@
     \item It would \textbf{only} detect fair queuing at \textbf{flow startup}. 
-    But if the bottleneck link changes during a flow, it could be that the previous bottleneck had fair queuing while the new one doesn't. This wouldn't have been detected. 
+    But if the bottleneck link changes during a flow, it could be the case that the previous bottleneck had fair queuing while the new one doesn't. This wouldn't have been detected. 
     \item It would detect fair queuing only after filling the queue at the bottleneck completely, \textbf{causing packet loss}. 
@@
 
-We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq} and \textit{fq\_codel} for fair queuing. 
+We emulate the network using \textit{mininet} and use Linux's \textit{pfifo} queuing discipline in the case there is no fair queuing and \textit{fq\_codel} and \textit{fq} for fair queuing. 
 For the evaluation we choose a simple network in which the sender is connected to the receiver via a switch.
@@
 
-We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy for both fq and fq\_codel, although experiments showed that accuracy depends on the configuration of the queue manager,
+We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy for both fq\_codel and fq, although experiments showed that accuracy depends on the configuration of the queue manager,
 the link capacity, base delay etc.
 Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cake} \cite{hoiland-jorgensen_piece_2018}, \textit{PIE} \cite{pan_pie_2013}, \textit{cocoa} \cite{bachl_cocoa_2019} or \textit{LFQ} \cite{bachl_lfq_2020}.
+
+If one were to deploy Tonopah on the internet, one could need to implement an API, through which applications can choose how much delay they allow, in case fair queuing is detected.
 

### 2022-10-06 tex/paper.tex
Fixed figures

@@
 If fair queuing is detected, the congestion control would not have to take cross traffic into account, which allows for more flexibility. 
-In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path, with very high accuracy. 
+In this paper, we develop the first algorithm that continuously checks if fair queuing is present on a path, with an accuracy of over 95\%. 
 When fair queuing is detected, a different congestion control can be chosen, which can result in reduced latency.
@@
 \subfloat[\footnotesize Queuing delay in case of \textbf{fair queuing}.]{
-\includegraphics[width=\columnwidth]{{"../sample/plots/concept_fq"}.pdf}
+\includegraphics[width=\columnwidth]{{"figures/concept_fq"}.pdf}
 \label{fig:fq_detail}}
@@
 \subfloat[\footnotesize Queuing delay in case there is \textbf{no fair queuing}.]{
-\includegraphics[width=\columnwidth]{{"../sample/plots/concept_fcfs"}.pdf}
+\includegraphics[width=\columnwidth]{{"figures/concept_fcfs"}.pdf}
 \label{fig:no_fq_detail}}
@@
 When different applications send packets over the internet, one application can send more than the other and thus unfairly take a larger share of bandwidth.
-Also, a large queue can form if one flow sends more than it should.
+Also, a queue can form if one flow sends more than it should.
 This can result in unfairness and bad user experience.
@@
 The dominant subflow receives $\frac{2}{3}$ of the total allowed sending rate, while the non-dominant one receives the remaining $\frac{1}{3}$.
+However, this can be changed as long as it is made sure that one subflow gets a significantly larger share than the other one.
 
@@
 In the following evaluations we always run each network flow for 90\,s. Tonopah measures whether there is fair queuing or not continuously. 
-If in the evaluation scenario there is fair queuing and Tonopah detects fair queuing correctly for 80\,s and fails to detect it for 10\,s, we 
+If there is fair queuing in the evaluation scenario and Tonopah detects fair queuing correctly for 80\,s and fails to detect it for 10\,s, we 
 state the accuracy as $\frac{80}{90} = 89\%$. 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_all"}.pdf}
+\includegraphics[width=\columnwidth]{{"figures/results_pfifo_new_tonopah_all"}.pdf}
 \caption{Detection accuracy in case there's \textbf{no fair queuing} (pfifo). The overall accuracy is 99.4\%.}
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah_all"}.pdf}
+\includegraphics[width=\columnwidth]{{"figures/results_fq_codel_new_tonopah_all"}.pdf}
 \caption{Detection accuracy in case there is \textbf{fair queuing} (fq\_codel). The overall accuracy is 95.8\%.}
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_new_tonopah_all"}.pdf}
+\includegraphics[width=\columnwidth]{{"figures/results_fq_new_tonopah_all"}.pdf}
 \caption{Detection accuracy in case there is \textbf{fair queuing} (fq). The overall accuracy is 95.4\%.}
@@
 \centering
-\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah"}.pdf}
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"figures/results_pfifo_new_tonopah"}.pdf}
 \label{fig:ecdf_pfifo}}
-\subfloat[\footnotesize fq\_codel]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_codel_new_tonopah"}.pdf}
+\subfloat[\footnotesize fq\_codel]{\includegraphics[width=0.33\columnwidth]{{"figures/results_fq_codel_new_tonopah"}.pdf}
 \label{fig:ecdf_fq_codel}}
-\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah"}.pdf}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"figures/results_fq_new_tonopah"}.pdf}
 \label{fig:ecdf_fq}}
@@
 
+\subsection{Cross-traffic}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{{"figures/results_pfifo_new_tonopah_iperf_all"}.pdf}
+\caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}. The overall accuracy is 99.9\%.}
+\label{fig:no_fq_crosstraffic}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{{"figures/results_fq_new_tonopah_iperf_all"}.pdf}
+\caption{Detection accuracy in case there is \textbf{fair queuing} (fq) under the presence of \textbf{cross-traffic}. The overall accuracy is 98.1\%.}
+\label{fig:fq_crosstraffic}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"figures/results_pfifo_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_pfifo}}
+\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"figures/results_fq_new_tonopah_iperf"}.pdf}
+\label{fig:ecdf_crosstraffic_fq}}
+\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing when there is \textbf{cross-traffic}.}
+\label{fig:ecdf_crosstraffic}
+\end{figure}
+
+To evaluate Tonopah's performance under cross-traffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
+\autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic, Tonpah's detection accuracy is virtually unchanged or even slightly better.
+We tried to use fq\_codel as in \autoref{fig:fq} but we encountered problems when using fq\_codel in combination with iperf3.
+Frequently, iperf3 would lose all packets, have a timeout after a couple of seconds and stop.
+Since we could not resolve this issue, we used fq instead of fq\_codel, for which this issue didn't occur.
+
+\autoref{fig:ecdf_crosstraffic} shows the empirical cumulative distribution functions when there is cross-traffic.
+
 \subsection{Throughput and Queuing Delay of Tonopah}
 
-\begin{table}
+\begin{table}[h]
 \begin{minipage}{\columnwidth}
@@
 
-\subsection{Cross-traffic}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf_all"}.pdf}
-\caption{Detection accuracy in case there's \textbf{no fair queuing} under the presence of \textbf{cross-traffic}. The overall accuracy is 99.9\%.}
-\label{fig:no_fq_crosstraffic}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf_all"}.pdf}
-\caption{Detection accuracy in case there is \textbf{fair queuing} (fq) under the presence of \textbf{cross-traffic}. The overall accuracy is 98.1\%.}
-\label{fig:fq_crosstraffic}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\subfloat[\footnotesize pfifo]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_pfifo_new_tonopah_iperf"}.pdf}
-\label{fig:ecdf_crosstraffic_pfifo}}
-\subfloat[\footnotesize fq]{\includegraphics[width=0.33\columnwidth]{{"../sample/plots/results_fq_new_tonopah_iperf"}.pdf}
-\label{fig:ecdf_crosstraffic_fq}}
-\caption{Empirical cumulative distribution functions for detection of the abscence/presence of fair queuing when there is \textbf{cross-traffic}.}
-\label{fig:ecdf_crosstraffic}
-\end{figure}
-
-To evaluate Tonopah's performance under cross-traffic, we use iperf3 with the NewReno congestion control. We start iperf 4\,s before Tonopah so that it has time to saturate the link. 
-\autoref{fig:no_fq_crosstraffic} and \autoref{fig:fq_crosstraffic} show that also in the presence of cross-traffic, Tonpah's detection accuracy is virtually unchanged or even slightly better.
-We tried to use fq\_codel as in \autoref{fig:fq} but we encountered problems when using fq\_codel in combination with iperf3.
-Frequently, iperf3 would lose all packets, have a timeout after a couple of seconds and stop.
-Since we could not resolve this issue, we used fq instead of fq\_codel, for which this issue didn't occur.
-
-\autoref{fig:ecdf_crosstraffic} shows the empirical cumulative distribution functions when there is cross-traffic.
-
 \section{Discussion}
@@
 
-If one were to deploy Tonopah on the internet, one could need to implement an API, through which applications can choose how much delay they allow, in case fair queuing is detected.
+If one were to deploy Tonopah on the internet, one could implement an API, through which applications can choose how much delay they allow, in case fair queuing is detected.
 

### 2023-05-28 picoquic/new_tonopah.c
Changes

@@ -667,7 +667,7 @@ static void picoquic_new_tonopah_delete(picoquic_path_t* path_x)
         path_x->congestion_alg_state = NULL;
         deleted_paths += 1;
     }
-    if (deleted_paths >= 2 && s_port == 4433) {
+    if (deleted_paths >= 2 && (s_port == 4433 || s_port == 4434)) {
         puts("The tonopah server (this is a server, right?) doesn't support several consecutive connections at the moment, exiting");
         exit(0);
     }

### 2023-05-28 sample/network_sim.py

@@ -136,7 +136,7 @@ for delay in (10, 50, 100):
 
                 strings = (
                     f"tc qdisc del dev {opt.interface} root", 
-                    f"tc qdisc add dev {opt.interface} root handle 1: netem{f' delay {int(opt.delay/2) if with_delay else 0}ms'}", 
+                    f"tc qdisc add dev {opt.interface} root handle 1: netem{f' delay {int(opt.delay/2) if with_delay else 0}ms'} limit 1000000", 
                     f"tc qdisc add dev {opt.interface} parent 1: handle 2: htb default 21", 
                     f"tc class add dev {opt.interface} parent 2: classid 2:21 htb rate {opt.rate if with_delay else 1000}mbit", 
                     f"tc qdisc add dev {opt.interface} parent 2:21 handle 3: {qdisc_string}"

### 2023-05-28 tex/paper.tex
Add section to tex about splitting data according to priority

@@
 
-We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy for both fq\_codel and fq, although experiments showed that accuracy depends on the configuration of the queue manager,
-the link capacity, base delay etc.
+We showed that continuous fair queuing detection is feasible and can achieve a high detection accuracy for both fq\_codel and fq, although experiments showed that accuracy depends on the configuration of the queue manager, the link capacity, base delay etc.
 Further experiments could be done to finetune Tonopah and also experiment with other queue managers such as \textit{cake} \cite{hoiland-jorgensen_piece_2018}, \textit{PIE} \cite{pan_pie_2013}, \textit{cocoa} \cite{bachl_cocoa_2019} or \textit{LFQ} \cite{bachl_lfq_2020}.
@@
 If one were to deploy Tonopah on the internet, one could implement an API, through which applications can choose how much delay they allow, in case fair queuing is detected.
+
+Since the subflow with the lower sending rate might experience lower queuing delay in case of fair queuing, it might be beneficial to include more important in this subflow while the less urgent data can be sent in the subflow with the higher sending rate.
+This might, for example, be useful for cloud gaming: One could separate the video data into two parts: One part which includes a low resolution base video, which is transferred with in the subflow with low queuing delay, while additional data, which increases the video quality, can be transferred in the subflow that might experience higher queuing delay.
 

### 2023-06-09 tex/paper.tex
Typos

--- 2022-10-06 tex/paper.tex
@@
 If one were to deploy Tonopah on the internet, one could implement an API, through which applications can choose how much delay they allow, in case fair queuing is detected.
+
+Since the subflow with the lower sending rate might experience lower queuing delay in case of fair queuing, it might be beneficial to include more important data in this subflow while the less urgent data can be sent in the subflow with the higher sending rate.
+This might, for example, be useful for cloud gaming: One could separate the video data into two parts: One part which includes a low resolution base video, which is transferred in the subflow with low queuing delay, while additional data, which increase the video quality, can be transferred in the subflow that might experience higher queuing delay.
 
